### 2024-03-03 outputs/paper/paper.qmd
Data and Graphs added

---
title: "childlessness-in-america"
author: Mehrnoush Mohammadi
thanks: "Code and data are available at: https://github.com/mehrnoush68/childlessness-in-america.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(knitr)
library(kableExtra)
library(here)
library(ggplot2)
library(formattable)

```

# Introduction

# Data {#sec-data}

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: nochild_age_graph

table_nochild_age <-
  read_csv(
    file = "../data/clean_data_age.csv",
    show_col_types = FALSE
  )

nochild_age_data_long <- tidyr::pivot_longer(table_nochild_age, -age, names_to = "year", values_to = "percentage")

ggplot(nochild_age_data_long, aes(x = as.integer(year), y = percentage, color = age)) +
  geom_line() +
  labs(x = "Year", y = "Percentage", title = "Percentage over Time by Age Group", color = "Age Group") +
  theme_minimal()

```

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: nochild_class_graph

table_nochild_class <-
  read_csv(
    file = "../data/clean_data_class.csv",
    show_col_types = FALSE
  )

# Pivot the data from wide to long format
data_long <- pivot_longer(table_nochild_class, -`Class Identification`, names_to = "Year", values_to = "Percentage")

# Convert Year column to integer
data_long$Year <- as.integer(data_long$Year)

# Plot the data
ggplot(data_long, aes(x = Year, y = Percentage, color = `Class Identification`)) +
  geom_line() +
  labs(x = "Year", y = "Percentage", title = "Percentage over Time by Class", color = "Breakdown Category") +
  theme_minimal()

```

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: nochild_degree_graph

table_nochild_degree <-
  read_csv(
    file = "../data/clean_data_degree.csv",
    show_col_types = FALSE
  )

# Pivot the data from wide to long format
data_long <- pivot_longer(table_nochild_degree, -`Degree`, names_to = "Year", values_to = "Percentage")

# Convert Year column to integer
data_long$Year <- as.integer(data_long$Year)

# Plot the data
ggplot(data_long, aes(x = Year, y = Percentage, color = `Degree`)) +
  geom_line() +
  labs(x = "Year", y = "Percentage", title = "Percentage over Time by Higest Degree", color = "Degree") +
  theme_minimal()

```

```{r}
#| echo: false
#| include: true
#| warning: false
#| label: nochild_gender_graph

table_nochild_sex <-
  read_csv(
    file = "../data/clean_data_sex.csv",
    show_col_types = FALSE
  )

# Pivot the data from wide to long format
data_long <- pivot_longer(table_nochild_sex, -`Gender`, names_to = "Year", values_to = "Percentage")

# Convert Year column to integer
data_long$Year <- as.integer(data_long$Year)

# Plot the data
ggplot(data_long, aes(x = Year, y = Percentage, color = `Gender`)) +
  geom_line() +
  labs(x = "Year", y = "Percentage", title = "Percentage over Time by Gender", color = "Gender") +
  theme_minimal()

```

# Results

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# References

### 2024-03-03 scripts/00-simulate_data.R

@@ -1,19 +1,17 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates data for the data sets no child, age, class identification, degree, and gender
+# Author: Mehrnoush Mohammadi
+# Date: 2 March 2024
+# Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(janitor)
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+set.seed(1)
 
 
 

### 2024-03-03 scripts/01-download_data.R

@@ -1,26 +1,37 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads the data from US General Social Survey
+# Author: Mehrnoush Mohammadi
+# Date: 2 March 2024
+# Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Download the raw data from US General Social Survey (https://gssdataexplorer.norc.org/trends)
 
 
 #### Workspace setup ####
-library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
 
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+#Note: data is behind sign-in
+# data can be accessed at https://gssdataexplorer.norc.org/trends from the US General Social Survey
 
-
-
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
-
+#### Load data ####
+nochild_age_data <-
+  read_csv(
+    file = "inputs/data/nochild_age.csv",
+    show_col_types = FALSE,
+  )
+nochild_class_data <-
+  read_csv(
+    file = "inputs/data/nochild_class.csv",
+    show_col_types = FALSE,
+  )
+nochild_sex_data <-
+  read_csv(
+    file = "inputs/data/nochild_sex.csv",
+    show_col_types = FALSE,
+  )
+nochild_degree_data <-
+  read_csv(
+    file = "inputs/data/nochild_higestdegree.csv",
+    show_col_types = FALSE,
+  )
          

### 2024-03-03 scripts/02-data_cleaning.R

@@ -1,44 +1,94 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw plane data childlessness in America
+# Author: Mehrnoush Mohammadi
+# Date: 2 March 2024
+# Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: need to have run 01-download_data.R
+
 
 #### Workspace setup ####
 library(tidyverse)
+library(readr)
+library(dplyr)
+library(knitr)
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+#### Load data ####
+nochild_age_data <-
+  read_csv(
+    file = "inputs/data/nochild_age.csv",
+    show_col_types = FALSE,
+  )
+nochild_class_data <-
+  read_csv(
+    file = "inputs/data/nochild_class.csv",
+    show_col_types = FALSE,
+  )
+nochild_sex_data <-
+  read_csv(
+    file = "inputs/data/nochild_sex.csv",
+    show_col_types = FALSE,
+  )
+nochild_degree_data <-
+  read_csv(
+    file = "inputs/data/nochild_higestdegree.csv",
+    show_col_types = FALSE,
+  )
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
[…]

### 2024-03-04 scripts/00-simulate_data.R
Simulate and test data added

@@ -8,10 +8,47 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(janitor)
 
 #### Simulate data ####
-set.seed(1)
+# Simulated data for table_nochild_age
+set.seed(123)  # For reproducibility
+age_groups <- c("18-34", "35-49", "50-64", "65+")
+years <- 1972:2022
 
+# Generate random percentages for each age group and year
+percentages_age <- matrix(runif(length(age_groups) * length(years), min = 5, max = 70), nrow = length(age_groups))
+colnames(percentages_age) <- years
+rownames(percentages_age) <- age_groups
 
+print(percentages_age)
+
+# Simulated data for table_nochild_class
+class_labels <- c("Lower class", "Working class", "Middle class", "Upper class")
+
+# Generate random percentages for each class and year
+percentages_class <- matrix(runif(length(class_labels) * length(years), min = 5, max = 70), nrow = length(class_labels))
+colnames(percentages_class) <- years
+rownames(percentages_class) <- class_labels
+
+print(percentages_class)
+
+# Simulated data for table_nochild_degree
+degree_labels <- c("High school diploma", "Bachelor's degree", "Master's degree", "Doctoral degree")
+
+# Generate random percentages for each degree and year
+percentages_degree <- matrix(runif(length(degree_labels) * length(years), min = 5, max = 70), nrow = length(degree_labels))
+colnames(percentages_degree) <- years
+rownames(percentages_degree) <- degree_labels
+
+print(percentages_degree)
+
+# Simulated data for table_nochild_sex
+gender_labels <- c("Male", "Female")
+
+# Generate random percentages for each gender and year
+percentages_sex <- matrix(runif(length(gender_labels) * length(years), min = 5, max = 70), nrow = length(gender_labels))
+colnames(percentages_sex) <- years
+rownames(percentages_sex) <- gender_labels
+
+print(percentages_sex)
 

### 2024-03-04 scripts/01-download_data.R

@@ -10,8 +10,9 @@
 #### Workspace setup ####
 library(tidyverse)
 
-#Note: data is behind sign-in
+# Note: data is behind sign-in
 # data can be accessed at https://gssdataexplorer.norc.org/trends from the US General Social Survey
+# data was downloaded as .xlsx file format and saved as .csv format on the computer
 
 #### Load data ####
 nochild_age_data <-

### 2024-03-04 scripts/02-data_cleaning.R

@@ -60,6 +60,10 @@ clean_data_degree <- function(x) {
 clean_data_age <- data_cleaned_age |>
   mutate(across(-1, clean_data))
 
+# Rename the 'age' column to 'Age'
+clean_data_age <- clean_data_age |>
+  rename(Age = age)
+
 clean_data_class <- data_cleaned_class |>
   mutate(across(-1, clean_data))
 

### 2024-03-04 scripts/03-test_data.R

@@ -1,15 +1,106 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests the three data sets
+# Author: Mehrnoush Mohammadi
+# Date: 2 March 2024
+# Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites:
+  # 01-download_data.R
+  # 02-data_cleaning.R
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(here)
+
+#### Read in cleaned data sets ####
+table_nochild_age <-
+  read_csv(
+    file = "outputs/data/clean_data_age.csv",
+    show_col_types = FALSE
+  )
+
+table_nochild_class <-
+  read_csv(
+    file = "outputs/data/clean_data_class.csv",
+    show_col_types = FALSE
+  )
+
+table_nochild_degree <-
+  read_csv(
+    file = "outputs/data/clean_data_degree.csv",
+    show_col_types = FALSE
+  )
+
+table_nochild_sex <-
+  read_csv(
+    file = "outputs/data/clean_data_sex.csv",
+    show_col_types = FALSE
+  )
+
+# Convert the dataset from wide to long format
+
+table_nochild_age_long <- pivot_longer(table_nochild_age, -Age, names_to = "Year", values_to = "Percentage")
+table_nochild_class_long <- pivot_longer(table_nochild_class, -`Class Identification`, names_to = "Year", values_to = "Percentage")
+table_nochild_degree_long <- pivot_longer(table_nochild_degree, -`Degree`, names_to = "Year", values_to = "Percentage")
+table_nochild_sex_long <- pivot_longer(table_nochild_sex, -`Gender`, names_to = "Year", values_to = "Percentage")
+
+# Add year labels
+
+year_labels <- c(1972:2022)
+
+table_nochild_age_long$Year <- factor(table_nochild_age_long$Year, levels = as.numeric(year_labels)) 
+table_nochild_class_long$Year <- factor(table_nochild_class_long$Year, levels = as.numeric(year_labels)) 
[…]

### 2024-03-04 outputs/paper/paper.qmd
Paper analysis and Usage added

@@
 ---
-title: "childlessness-in-america"
+title: "Understanding the Landscape of Childlessness in the U.S.: A Data-Driven Exploration (50-Year Trends)"
 author: Mehrnoush Mohammadi
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper investigates trends in childlessness in the United States from 1970 to 2022.
+Data from the General Social Survey (GSS) is analyzed to consider factors like age, gender, education, class, and number of children.
+The analysis reveals a consistent rise in childlessness across various demographics, potentially linked to economic considerations, educational attainment, and shifting social norms.
+The paper explores the potential social and economic implications of this trend, highlighting the need for further research to navigate the complexities surrounding childlessness in today's society."
 format: pdf
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 
+#### Workspace setup ####
 library(tidyverse)
-library(palmerpenguins)
 library(knitr)
 library(kableExtra)
-library(here)
 library(ggplot2)
-library(formattable)
+library(dplyr)
 
-```
-
-# Introduction
-
-# Data {#sec-data}
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: nochild_age_graph
-
+#### Load data ####
 table_nochild_age <-
@@
 
-nochild_age_data_long <- tidyr::pivot_longer(table_nochild_age, -age, names_to = "year", values_to = "percentage")
+table_nochild_class <-
+  read_csv(
+    file = "../data/clean_data_class.csv",
+    show_col_types = FALSE
+  )
 
-ggplot(nochild_age_data_long, aes(x = as.integer(year), y = percentage, color = age)) +
-  geom_line() +
-  labs(x = "Year", y = "Percentage", title = "Percentage over Time by Age Group", color = "Age Group") +
-  theme_minimal()
+table_nochild_degree <-
+  read_csv(
+    file = "../data/clean_data_degree.csv",
+    show_col_types = FALSE
+  )
+
+table_nochild_sex <-
+  read_csv(
+    file = "../data/clean_data_sex.csv",
+    show_col_types = FALSE
+  )
 
 ```
+
+# Introduction
+The American family landscape is undergoing a significant transformation, marked by a growing trend of childlessness.
+Traditionally viewed as a societal norm, having children is now a conscious choice faced by an increasing number of individuals.
+This paper delves into this phenomenon by examining various datasets encompassing critical factors like age, gender, education, socioeconomic class, and national childbearing rates over the past 50 years.
+
+Analyzing these factors offers a window into understanding not only the prevalence of childlessness within different segments of the population but also the potential drivers behind this trend.
+Existing research suggests a complex interplay of factors, including economic considerations, rising educational attainment, changing social norms related to career and family life, and individual preferences.
+
+This paper aims to unpack these potential contributing factors and explore their impact on the rise of childlessness in America.
+By analyzing the data, I seek to answer crucial questions: What are the key factors contributing to this trend?
+How does childlessness manifest differently across various demographic groups?
+Finally, what are the potential social, economic, and individual consequences of this evolving family structure in the United States?
+Understanding these dynamics is crucial for shaping informed policy discussions and navigating the implications of childlessness in a rapidly changing world.
+
+The graphs and tables in this paper were created in R Studio using R [@citeR] and the analysis in a Quarto document.
+The analysis was conducted with the use of the `ggplot` [@ggplotRef], `tiddyverse` [@tidyverseRef], `knitr` [@knitrRef], `readr` [@readRef], `kableExtra` [@kableextraRef], `janitor` [@janitorRef] and `dyplr` [@dplRef ] packages.
+
+# Data
+This paper utilizes a multifaceted approach, drawing on various datasets to illuminate the complexities of childlessness in the United States.
+One primary source is the General Social Survey (GSS), a long-running and highly respected survey conducted by the National Opinion Research Center (NORC) at the University of Chicago [@gss], from 1970 to 2022.
+
+## Source Data
+
+The 2022 iteration of the GSS, for the first time in its history, incorporated a combination of in-person, web self-administered, and phone interview methods.
+This expansion in data collection methods offers a potentially richer and more diverse sample, while acknowledging that specific variables might be limited to certain modes.
+The GSS team is transparent about these limitations and welcomes feedback from the research community to continuously improve their methodology.
+
+From the GSS, I utilize several key variables relevant to my analysis:
+
+* Children: This variable captures the number of children ever born alive to the respondent, providing a direct measure of childlessness (having no children).
+* Year: This variable allows to analyze trends in childlessness over time, encompassing the chosen time frame for the survey.
+* Age: This variable, categorized into relevant age groups (e.g., 18-34, 35-49, etc.), enables us to understand how childlessness manifests across different age demographics.
+It is important to note that the GSS data underwent specific recoding procedures for this analysis.
+For instance, the original "CHILD" variable was recoded to create the "childs" variable used here, and age categories were grouped for visualization purposes.
+
+@tbl-nochild-age demonstrates a clear trend of increasing childlessness across age groups over the selected years.
+As evident from the data, the percentage of individuals without children has steadily risen in all age groups, with the most significant increase observed in the younger age groups (18-34 and 35-49).
+This trend suggests a shift in childbearing patterns, with individuals potentially delaying or forgoing parenthood altogether.
 
@@
 #| warning: false
-#| label: nochild_class_graph
+#| label: tbl-nochild-age
 
-table_nochild_class <-
-  read_csv(
-    file = "../data/clean_data_class.csv",
-    show_col_types = FALSE
-  )
+# Selected years to show better on the paper
+random_years <- c("Age", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_age_subset <- table_nochild_age |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_age_subset, 
+      caption = "Childless Percentage of Population by Age Group and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
+### Exploring Socioeconomic Class and Educational Attainment
+To gain a deeper understanding of the factors influencing these trends, I delve into additional data sets from NORC (@gss) examining the relationship between childlessness and socioeconomic class @tbl-nochild-class and educational attainment @tbl-nochild-degree.
+
+From the GSS, I utilize several key variables to enrich my analysis and explore the potential influence of various factors on childbearing decisions:
+
+* Socioeconomic Class: Data on socioeconomic class, often measured by income or occupation, will be utilized to understand the impact of economic factors on childbearing choices.
+* Educational Attainment: Data on educational attainment, such as highest degree obtained, will be incorporated to explore the relationship between educational level and childbearing.
+This data can be obtained from sources like the National Center for Education Statistics.
+
+@tbl-nochild-class reveals a potential link between socioeconomic class and childlessness.
+The data suggests that individuals in the lower class and working class exhibit higher percentages of childlessness compared to those in the middle class and upper class.
+This pattern might be attributed to various factors, such as financial constraints, job insecurity, or limited access to childcare resources experienced by individuals in lower socioeconomic classes.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-class
+
+# Selected years to show better on the paper
+random_years <- c("Class Identification", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_class_subset <- table_nochild_class |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_class_subset, 
+      caption = "Childless Percentage of Population by Class Indentification and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
+@tbl-nochild-degree explores the association between educational attainment and childlessness.
+Interestingly, the data indicates a nuanced relationship.
+While individuals with higher educational degrees (college+) display a slightly higher percentage of childlessness compared to those with lower degrees, the difference is not as substantial as observed with socioeconomic class.
+This suggests that educational attainment might play a complex role in childbearing decisions, potentially influenced by factors like career aspirations, financial considerations, and personal preferences.
+
+\newpage
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-degree
+
+# Selected years to show better on the paper
+random_years <- c("Degree", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_degree_subset <- table_nochild_degree |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_degree_subset, 
+      caption = "Childless Percentage of Population by Degree and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
+### Exploring Gender Differences
+It is crucial to acknowledge potential gender disparities in childbearing experiences.
+A separate analysis examining childlessness by gender (presented in @tbl-nochild-sex) aims to provide a comprehensive understanding by considering the unique experiences of men and women regarding childlessness.
+Data on gender is included from NORC [@gss] to allow for this investigation.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-sex
+
+# Selected years to show better on the paper
+random_years <- c("Gender", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_sex_subset <- table_nochild_sex |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_sex_subset, 
+      caption = "Childless Percentage of Population by Gender and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
+## Data Limitations
+While the data utilized in this paper provides valuable insights into childlessness in the United States, it is important to acknowledge some limitations:
+
+### Incomplete Data for Certain Years
+While the General Social Survey provides valuable data on childlessness trends, there are gaps in data availability for certain years.
+This creates a limitation in our ability to present a completely continuous picture of trends across the entire 50-year time frame.
+
+### Generalizability of the GSS
+
+The General Social Survey (GSS) is a valuable resource but may not be entirely representative of the entire U.S.
+population.
+Sampling bias and potential non-response bias inherent in any survey can limit the generalization of findings to the broader population.
+
+### Potential Measurement Errors
+
+Survey data may be susceptible to measurement errors due to various factors like social desirability bias, memory lapses, or misinterpretations of questions.
+While the GSS implements measures to minimize these errors, they cannot be entirely eliminated.
+
+### Limited Consideration of Gender Disparities
+
+While acknowledging the importance of gender disparities, the analysis presented here only provides a preliminary exploration through a single table (@tbl-nochild-sex).
+Further research dedicated to understanding nuanced gender differences in childbearing experiences and the complex interplay of individual and societal factors is needed.
+
+# Result and Analysis
+
+@fig-nochild-age illustrates of @tbl-nochild-age which is a consistent rise in childlessness across all age groups in the United States from 1970 to 2022.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-age
+#| fig-cap: "Percentage of Childless Adults by Age Group in the US (1970-2020)"
+
+nochild_age_data_long <- tidyr::pivot_longer(table_nochild_age, -Age, names_to = "year", values_to = "percentage")
+
+ggplot(nochild_age_data_long, aes(x = as.integer(year), y = percentage, color = Age)) +
+  geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "solid", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Age Group", color = "Age Group") +
+  theme_minimal()
+```
+
+\newpage
+
+It is important to acknowledge, however, that the data series has gaps for several years within this timeframe (1979, 1981, 1992, 1995, 1997, 1999, 2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017, and 2019).
+These missing data points can potentially limit the generalizability of the observed trends and hinder a comprehensive understanding of the childlessness over time.
+Despite these limitations, the available data reveals a steady increase in the percentage of individuals without children across all age groups.
+Notably, the most significant rise is observed among younger adults (18-34 and 35-49).
+As an example, the percentage of childless individuals in the 18-34 age group nearly doubled from 10% in 1970 to 19% in 2020.
+
+@fig-nochild-class visualizes @tbl-nochild-class which depicts the gradual increase in the percentage of childless individuals across three social classes (Working Class, Middle Class, Upper Class) in the United States from 1970 to 2020.
+The graph reveals an upward trend in childlessness for all classes over the five decades.
+
+While all classes show an increase, the data suggests the most significant rise occurred among the Working Class, with their percentage of childless individuals nearly doubling from 20% in 1970 to 40% in 2020.
+The Middle Class and Upper Class also experienced steady increases, though to a lesser extent, reaching around 32% and 25% childlessness, respectively, in 2020.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-class
+#| fig-cap: "Percentage Childless over Time by Class"
 
@@
   geom_line() +
-  labs(x = "Year", y = "Percentage", title = "Percentage over Time by Class", color = "Breakdown Category") +
+  geom_smooth(method = "loess", se = FALSE, linetype = "solid", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Class", color = "Breakdown Category") +
   theme_minimal()
@@
 ```
+
+@fig-nochild-degree visualization of @tbl-nochild-degree, presents a comparative analysis of childlessness across individuals with different levels of educational attainment (High School Diploma, Bachelor's Degree, Master's Degree or Higher) in the United States from 1970 to 2020.
+The graph utilizes line trends to depict the percentage of childless individuals within each degree category over the five decades.
+
+The data reveals a distinct pattern across educational groups.
+Individuals with a High School Diploma consistently exhibit the highest percentage of childlessness throughout the entire period, ranging from around 40% in 1970 to nearly 55% in 2020.
+Conversely, individuals with a Master's Degree or Higher consistently exhibit the lowest percentage of childlessness, starting at around 15% in 1970 and reaching approximately 25% in 2020.
+The trend for individuals with a Bachelor's Degree falls between the other two groups, displaying a rise from 25% in 1970 to around 35% in 2020.
 
@@
 #| warning: false
-#| label: nochild_degree_graph
-
-table_nochild_degree <-
-  read_csv(
-    file = "../data/clean_data_degree.csv",
-    show_col_types = FALSE
-  )
+#| label: fig-nochild-degree
+#| fig-cap: "Percentage Childless over Time by Level of Education"
 
@@
   geom_line() +
-  labs(x = "Year", y = "Percentage", title = "Percentage over Time by Higest Degree", color = "Degree") +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Degree", color = "Degree") +
   theme_minimal()
@@
 ```
+
+@fig-nochild-gender illustrates of @tbl-nochild-sex, explores gender differences in childlessness within the United States over a 50-year period (1970-2020).
+The graph utilizes line trends to depict the percentage of childless individuals categorized by gender (Male and Female).
+
+The data reveals a gradual increase in childlessness for both genders over the past five decades.
+However, females consistently exhibit a lower percentage of childlessness compared to males throughout the entire period.
+While the gap appears to be slightly narrowing over time, females still have a lower percentage by around 10 percentage points in 2020 (approximately 35% for females and 45% for males).
 
@@
 #| warning: false
-#| label: nochild_gender_graph
-
-table_nochild_sex <-
-  read_csv(
-    file = "../data/clean_data_sex.csv",
-    show_col_types = FALSE
-  )
+#| label: fig-nochild-gender
+#| fig-cap: "Percentage Childless over Time by Gender"
 
@@
   geom_line() +
-  labs(x = "Year", y = "Percentage", title = "Percentage over Time by Gender", color = "Gender") +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Gender", color = "Gender") +
   theme_minimal()
@@
 
-# Results
+The figures presented in this paper were generated using data from the [@gss].
+These figures visually represent the key findings related to the trends and patterns in childlessness across different age, social classes, educational attainment levels, and genders in the United States over the past five decades (1970-2020).
+It is important to remember that these figures solely present descriptive data and do not establish causal relationships.
+The Discussion section will delve deeper into potential explanations for the observed trends and explore the broader implications of these findings.
 
@@
 
-## First discussion point {#sec-first-point}
+## Rising Childlessness: Patterns and Trends
+The findings presented in this paper reveal a consistent and concerning trend of rising childlessness across various demographics in the United States.
+As Figure 1 illustrates, the percentage of individuals without children has steadily increased across all age groups over the past 50 years, with the most significant rise observed among younger individuals (18-34 and 35-49).
 
-## Second discussion point
+## Potential Explanations for Rising Childlessness
 
-## Third discussion point
+This paper cannot definitively establish causal relationships, but the data suggests several potential factors contributing to rising childlessness, all supported by existing research.
+These include economic concerns like rising childcare costs, student loan debt, and economic uncertainty.
+Additionally, individuals with higher educational attainment, particularly women, might delay or forgo childbearing to pursue career opportunities.
+Furthermore, shifting social norms related to childbearing, family structures, and career aspirations, especially for women, may be contributing factors.
+Finally, the decision to remain childless can also be influenced by individual values, lifestyle choices, and personal circumstances.
+
+## Societal Implications of Rising Childlessness
+
+The rising trend of childlessness carries potential consequences across various societal spheres.
+A shrinking pool of younger workers could pose challenges to economic growth and strain social security systems.
+Additionally, the rise of childless families could have implications for social support networks and intergenerational relationships.
+While childlessness can be a personal choice for some individuals and may even lead to a fulfilling life, it could also be associated with feelings of isolation and a lack of social support for others.
+These potential consequences highlight the need for further research and discussion on how to navigate the social, economic, and individual complexities associated with rising childlessness.
 
@@
 
-Weaknesses and next steps should also be included.
+This study has some limitations that need to be acknowledged.
+Firstly, the study relies only on data from the GSS, which may not be entirely representative of the US population.
+Secondly, the cross-sectional design of the GSS data limits the ability to establish causal relationships between variables.
+Lastly, surveys are prone to biases such as social desirability bias or memory lapses, which may impact the accuracy of the findings.
+
+To gain a deeper understanding of childlessness, there are some crucial next steps that need to be taken.
+Firstly, employing longitudinal studies will allow the exploration of causal relationships between factors and childbearing decisions over time.
+Secondly, incorporating diverse data sources, such as qualitative interviews and other surveys with broader representation, can provide richer insights and explore the nuances of individual experiences.
+Lastly, utilizing a mixed methods approach, combining quantitative and qualitative data analysis, will offer a more comprehensive understanding of both the broader trends and the lived experiences of individuals facing childlessness.
+
+By addressing these limitations and pursuing the proposed next steps, future research can contribute significantly to a more nuanced and in-depth understanding of childlessness and its various aspects.
+
+# Conclusion
+This paper investigated the rising trend of childlessness in the United States, highlighting its prevalence across various demographics.
+By analyzing data from the GSS, it revealed consistent and concerning increases in the percentage of individuals without children.
+While definitive causal explanations remain elusive due to limitations in the analysis, the discussion explored potential contributing factors, including economic considerations, educational aspirations, shifting social norms, and individual preferences.
+Additionally, the potential consequences of this trend, spanning across social, economic, and individual realms, were examined.
+
+Moving forward, further research with longitudinal studies and diverse data sources including qualitative interviews, alongside a mixed methods approach, is crucial to gain a comprehensive understanding of childlessness, its nuances, and its multifaceted implications.
+This knowledge can inform future policies and interventions aimed at fostering a supportive environment for individuals and families as they navigate the complex decisions around childbearing in the 21st century.
 
@@
 
-# Appendix {-}
+# Appendix
+The survey questions are available here under each trend: https://gssdataexplorer.norc.org/trends
+
+## Survey Questions from GSS
+(@) How many children have you ever had?  Please count all that were born alive at any time (including any you had from a previous marriage).
+- 0
+- 1
+- 2
+- 3
+- 4
+- 5
+- 6
+- 7
+- 8 or more
+
+(@) RESPONDENT'S AGE
+(short answer text)
+
+(@) If you were asked to use one of four names for your social class, which would you say you belong in: 
+- The lower class 
+- The working class
+- The middle class
+- The upper class
+
+(@) RESPONDENT'S DEGREE:
+- College +
+- High school
+- Less than high school
+
+(@) RESPONDENT'S SEX:
+- Female
+- Male
 

### 2024-03-06 outputs/paper/paper.qmd
semifinal changes of the paper

@@
 ---
-title: "Understanding the Landscape of Childlessness in the U.S.: A Data-Driven Exploration (50-Year Trends)"
+title: "Understanding the Landscape of Childlessness in the U.S.: A Data-Driven Exploration (1972-2022)"
 author: Mehrnoush Mohammadi
@@
 date-format: long
-abstract: "This paper investigates trends in childlessness in the United States from 1970 to 2022.
+abstract: "This paper investigates trends in childlessness in the United States from 1972 to 2022.
 Data from the General Social Survey (GSS) is analyzed to consider factors like age, gender, education, class, and number of children.
@@
 This paper utilizes a multifaceted approach, drawing on various datasets to illuminate the complexities of childlessness in the United States.
-One primary source is the General Social Survey (GSS), a long-running and highly respected survey conducted by the National Opinion Research Center (NORC) at the University of Chicago [@gss], from 1970 to 2022.
+One primary source is the General Social Survey (GSS), a long-running and highly respected survey conducted by the National Opinion Research Center (NORC) at the University of Chicago [@gss], from 1972 to 2022.
 
@@
 
-@fig-nochild-age illustrates of @tbl-nochild-age which is a consistent rise in childlessness across all age groups in the United States from 1970 to 2022.
+@fig-nochild-age illustrates of @tbl-nochild-age which is a consistent rise in childlessness across all age groups in the United States from 1972 to 2022.
 
@@
 #| label: fig-nochild-age
-#| fig-cap: "Percentage of Childless Adults by Age Group in the US (1970-2020)"
+#| fig-cap: "Percentage of Childless Adults by Age Group in the US (1972-2020)"
 
@@
   theme_minimal()
+
+# Calculation of Percentage Increase in 18-34
+initial_value <- 46.07
+final_value <- 65.76
+percentage_increase <- ((final_value - initial_value) / initial_value) * 100
+
+#percentage_increase
 ```
@@
 It is important to acknowledge, however, that the data series has gaps for several years within this timeframe (1979, 1981, 1992, 1995, 1997, 1999, 2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017, and 2019).
-These missing data points can potentially limit the generalizability of the observed trends and hinder a comprehensive understanding of the childlessness over time.
+These missing data points can potentially limit the generalization of the observed trends and hinder a comprehensive understanding of the childlessness over time.
 Despite these limitations, the available data reveals a steady increase in the percentage of individuals without children across all age groups.
 Notably, the most significant rise is observed among younger adults (18-34 and 35-49).
-As an example, the percentage of childless individuals in the 18-34 age group nearly doubled from 10% in 1970 to 19% in 2020.
+As an example, the percentage of childless individuals in the 18-34 age group increased by approximately 42.70% from 1972 to 2022.
 
-@fig-nochild-class visualizes @tbl-nochild-class which depicts the gradual increase in the percentage of childless individuals across three social classes (Working Class, Middle Class, Upper Class) in the United States from 1970 to 2020.
+@fig-nochild-class visualizes @tbl-nochild-class which depicts the gradual increase in the percentage of childless individuals across four social classes (Lower Class, Working Class, Middle Class, Upper Class) in the United States from 1972 to 2022.
 The graph reveals an upward trend in childlessness for all classes over the five decades.
-
-While all classes show an increase, the data suggests the most significant rise occurred among the Working Class, with their percentage of childless individuals nearly doubling from 20% in 1970 to 40% in 2020.
-The Middle Class and Upper Class also experienced steady increases, though to a lesser extent, reaching around 32% and 25% childlessness, respectively, in 2020.
 
@@
 
-@fig-nochild-degree visualization of @tbl-nochild-degree, presents a comparative analysis of childlessness across individuals with different levels of educational attainment (High School Diploma, Bachelor's Degree, Master's Degree or Higher) in the United States from 1970 to 2020.
+People and households are divided into groups based on their socioeconomic status.
+The lower class experiences economic hardship, while the Working Class earns limited wages and faces job insecurity.
+The middle class has white-collar or professional occupations and enjoys a comfortable lifestyle.
+The upper class has substantial wealth and enjoys a privileged lifestyle with exclusive amenities [@pewresearch2022].
+
+While all classes show an increase in @fig-nochild-class, the data suggests the most significant rise occurred among the Lower Class, with their percentage of childless individuals nearly doubling from approximately 17.16% in 1972 to around 34.46% in 2022, marking a substantial increase of approximately 101.04% over the specified period.
+The Working Class also experienced a significant increase, from around 21.83% in 1972 to approximately 31.44% in 2022, indicating a rise of roughly 43.98%.
+Additionally, the Middle Class saw a slight increase, from approximately 28.24% in 1972 to about 30.09% in 2022, representing a rise of approximately 6.55%.
+Similarly, the Upper Class saw an increase, with their percentage of childless individuals climbing from approximately 20.24% in 1972 to about 27.87% in 2022, indicating an increase of roughly 37.71%.
+
+@fig-nochild-degree visualization of @tbl-nochild-degree, presents a comparative analysis of childlessness across individuals with different levels of educational attainment (Less than High School, High School and College's Degree or Higher) in the United States from 1972 to 2022.
 The graph utilizes line trends to depict the percentage of childless individuals within each degree category over the five decades.
@@
 The data reveals a distinct pattern across educational groups.
-Individuals with a High School Diploma consistently exhibit the highest percentage of childlessness throughout the entire period, ranging from around 40% in 1970 to nearly 55% in 2020.
-Conversely, individuals with a Master's Degree or Higher consistently exhibit the lowest percentage of childlessness, starting at around 15% in 1970 and reaching approximately 25% in 2020.
-The trend for individuals with a Bachelor's Degree falls between the other two groups, displaying a rise from 25% in 1970 to around 35% in 2020.
-
+Individuals with a high school diploma consistently exhibit the highest percentage of childlessness throughout the entire period, ranging from around 28.15% in 1972 to 34.56% in 2022.
+Conversely, The percentage of childless individuals with a College and higher degree has decreased from 32.85% in 1972 to 29.84% in 2022.
+The trend for individuals with a with less than a high school education increased from 16.97% in 1972 to 22.90% in 2022.
 ```{r}
@@
 
-@fig-nochild-gender illustrates of @tbl-nochild-sex, explores gender differences in childlessness within the United States over a 50-year period (1970-2020).
+\newpage
+
+@fig-nochild-gender illustrates of @tbl-nochild-sex, explores gender differences in childlessness within the United States over a 50-year period (1972-2022).
 The graph utilizes line trends to depict the percentage of childless individuals categorized by gender (Male and Female).
@@
 However, females consistently exhibit a lower percentage of childlessness compared to males throughout the entire period.
-While the gap appears to be slightly narrowing over time, females still have a lower percentage by around 10 percentage points in 2020 (approximately 35% for females and 45% for males).
+While the gap appears to be slightly narrowing over time, females still have a lower percentage by around 8 percentage points in 2022 (approximately 28.32% for females and 36.09% for males).
 
@@
 The figures presented in this paper were generated using data from the [@gss].
-These figures visually represent the key findings related to the trends and patterns in childlessness across different age, social classes, educational attainment levels, and genders in the United States over the past five decades (1970-2020).
+These figures visually represent the key findings related to the trends and patterns in childlessness across different age, social classes, educational attainment levels, and genders in the United States over the past five decades (1972-2022).
 It is important to remember that these figures solely present descriptive data and do not establish causal relationships.

### 2024-03-10 scripts/01-download_data.R
update datasets

@@ -9,6 +9,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(readr)
 
 # Note: data is behind sign-in
 # data can be accessed at https://gssdataexplorer.norc.org/trends from the US General Social Survey
@@ -35,4 +36,13 @@ nochild_degree_data <-
     file = "inputs/data/nochild_higestdegree.csv",
     show_col_types = FALSE,
   )
-         
+nochild_degree_data <-
+  read_csv(
+    file = "inputs/data/nochild_health.csv",
+    show_col_types = FALSE,
+  )
+nochild_degree_data <-
+  read_csv(
+    file = "inputs/data/nochild_marital.csv",
+    show_col_types = FALSE,
+  )

### 2024-03-10 scripts/00-simulate_data.R
new datasets on health and marital status

@@ -52,3 +52,23 @@ rownames(percentages_sex) <- gender_labels
 
 print(percentages_sex)
 
+# Simulated data for table_nochild_health
+health_labels <- c("Excellent", "Good", "Fair", "Poor")
+
+# Generate random percentages for each gender and year
+percentages_health <- matrix(runif(length(health_labels) * length(years), min = 5, max = 70), nrow = length(health_labels))
+colnames(percentages_health) <- years
+rownames(percentages_health) <- health_labels
+
+print(percentages_health)
+
+# Simulated data for table_nochild_marital
+marital_labels <- c("Never married", "Married", "Separated", "Divorced", "Widowed")
+
+# Generate random percentages for each gender and year
+percentages_marital <- matrix(runif(length(marital_labels) * length(years), min = 5, max = 70), nrow = length(marital_labels))
+colnames(percentages_marital) <- years
+rownames(percentages_marital) <- marital_labels
+
+print(percentages_marital)
+

### 2024-03-10 scripts/01-download_data.R

@@ -36,12 +36,12 @@ nochild_degree_data <-
     file = "inputs/data/nochild_higestdegree.csv",
     show_col_types = FALSE,
   )
-nochild_degree_data <-
+nochild_health_data <-
   read_csv(
     file = "inputs/data/nochild_health.csv",
     show_col_types = FALSE,
   )
-nochild_degree_data <-
+nochild_marital_data <-
   read_csv(
     file = "inputs/data/nochild_marital.csv",
     show_col_types = FALSE,

### 2024-03-10 scripts/02-data_cleaning.R

@@ -34,12 +34,25 @@ nochild_degree_data <-
     file = "inputs/data/nochild_higestdegree.csv",
     show_col_types = FALSE,
   )
+nochild_health_data <-
+  read_csv(
+    file = "inputs/data/nochild_health.csv",
+    show_col_types = FALSE,
+  )
+nochild_marital_data <-
+  read_csv(
+    file = "inputs/data/nochild_marital.csv",
+    show_col_types = FALSE,
+  )
 
 # Cleaning Data
 data_cleaned_age <- nochild_age_data
 data_cleaned_class <- nochild_class_data
 data_cleaned_sex <- nochild_sex_data
 data_cleaned_degree <- nochild_degree_data
+data_cleaned_health <- nochild_health_data
+data_cleaned_marital <- nochild_marital_data
+
 
 
 # Define a function to clean the data
@@ -55,23 +68,35 @@ clean_data_sex <- function(x) {
 clean_data_degree <- function(x) {
   as.numeric(gsub("\\s*\\(.*\\)", "", x))
 }
+clean_data_health <- function(x) {
+  as.numeric(gsub("\\s*\\(.*\\)", "", x))
+}
+clean_data_marital <- function(x) {
+  as.numeric(gsub("\\s*\\(.*\\)", "", x))
+}
 
 # Clean the data
 clean_data_age <- data_cleaned_age |>
-  mutate(across(-1, clean_data))
+  mutate(across(-1, clean_data_age))
 
 # Rename the 'age' column to 'Age'
 clean_data_age <- clean_data_age |>
   rename(Age = age)
 
 clean_data_class <- data_cleaned_class |>
-  mutate(across(-1, clean_data))
+  mutate(across(-1, clean_data_class))
 
 clean_data_sex <- data_cleaned_sex |>
-  mutate(across(-1, clean_data))
+  mutate(across(-1, clean_data_sex))
 
 clean_data_degree <- data_cleaned_degree |>
-  mutate(across(-1, clean_data))
+  mutate(across(-1, clean_data_degree))
+
+clean_data_marital <- data_cleaned_marital |>
+  mutate(across(-1, clean_data_marital))
+
+clean_data_health <- data_cleaned_health |>
+  mutate(across(-1, clean_data_health))
 
 
 # Save cleaned data
@@ -95,4 +120,14 @@ write_csv(
   file = "outputs/data/clean_data_degree.csv"
 )
 
+write_csv(
[…]

### 2024-03-10 scripts/02-data_cleaning.R
stimulate and test new datasets

@@ -125,7 +125,6 @@ write_csv(
   file = "outputs/data/clean_data_health.csv"
 )
 
-
 write_csv(
   x = clean_data_marital,
   file = "outputs/data/clean_data_marital.csv"

### 2024-03-10 scripts/03-test_data.R

@@ -38,12 +38,26 @@ table_nochild_sex <-
     show_col_types = FALSE
   )
 
+table_nochild_health <-
+  read_csv(
+    file = "outputs/data/clean_data_health.csv",
+    show_col_types = FALSE
+  )
+
+table_nochild_marital <-
+  read_csv(
+    file = "outputs/data/clean_data_marital.csv",
+    show_col_types = FALSE
+  )
+
 # Convert the dataset from wide to long format
 
 table_nochild_age_long <- pivot_longer(table_nochild_age, -Age, names_to = "Year", values_to = "Percentage")
 table_nochild_class_long <- pivot_longer(table_nochild_class, -`Class Identification`, names_to = "Year", values_to = "Percentage")
 table_nochild_degree_long <- pivot_longer(table_nochild_degree, -`Degree`, names_to = "Year", values_to = "Percentage")
 table_nochild_sex_long <- pivot_longer(table_nochild_sex, -`Gender`, names_to = "Year", values_to = "Percentage")
+table_nochild_health_long <- pivot_longer(table_nochild_health, -`Condition of health`, names_to = "Year", values_to = "Percentage")
+table_nochild_marital_long <- pivot_longer(table_nochild_marital, -`Marital status`, names_to = "Year", values_to = "Percentage")
 
 # Add year labels
 
@@ -53,11 +67,16 @@ table_nochild_age_long$Year <- factor(table_nochild_age_long$Year, levels = as.n
 table_nochild_class_long$Year <- factor(table_nochild_class_long$Year, levels = as.numeric(year_labels)) 
 table_nochild_degree_long$Year <- factor(table_nochild_degree_long$Year, levels = as.numeric(year_labels)) 
 table_nochild_sex_long$Year <- factor(table_nochild_sex_long$Year, levels = as.numeric(year_labels)) 
+table_nochild_marital_long$Year <- factor(table_nochild_marital_long$Year, levels = as.numeric(year_labels)) 
+table_nochild_health_long$Year <- factor(table_nochild_health_long$Year, levels = as.numeric(year_labels)) 
+
 
 print(table_nochild_age_long)
 print(table_nochild_class_long)
 print(table_nochild_degree_long)
 print(table_nochild_sex_long)
+print(table_nochild_health_long)
+print(table_nochild_marital_long)
 
 
 #### Test data ####
[…]

### 2024-03-11 outputs/paper/paper.qmd
Survey questions updated
new dataset updates

@@
 title: "Understanding the Landscape of Childlessness in the U.S.: A Data-Driven Exploration (1972-2022)"
-author: Mehrnoush Mohammadi
+author: Mehrnoush Mohammadi & Quang Mai
 thanks: "Code and data are available at: https://github.com/mehrnoush68/childlessness-in-america.git"
@@
 abstract: "This paper investigates trends in childlessness in the United States from 1972 to 2022.
-Data from the General Social Survey (GSS) is analyzed to consider factors like age, gender, education, class, and number of children.
+Data from the General Social Survey (GSS) is analyzed to consider factors like age, gender, education, class, health condition, marital status and number of children.
 The analysis reveals a consistent rise in childlessness across various demographics, potentially linked to economic considerations, educational attainment, and shifting social norms.
@@
 library(dplyr)
+library(readr)
 
@@
 
+table_nochild_marital <-
+  read_csv(
+    file = "../data/clean_data_marital.csv",
+    show_col_types = FALSE
+  )
+
+table_nochild_health <-
+  read_csv(
+    file = "../data/clean_data_health.csv",
+    show_col_types = FALSE
+  )
+
 ```
 
-# Introduction
+# Introduction {#sec-intro}
 The American family landscape is undergoing a significant transformation, marked by a growing trend of childlessness.
-Traditionally viewed as a societal norm, having children is now a conscious choice faced by an increasing number of individuals.
-This paper delves into this phenomenon by examining various datasets encompassing critical factors like age, gender, education, socioeconomic class, and national childbearing rates over the past 50 years.
+Traditionally viewed as a societal norm, having children now becomes a choice.
+This paper looks into this phenomenon by examining various datasets on factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they affect childbearing rates over the past 50 years.
 
 Analyzing these factors offers a window into understanding not only the prevalence of childlessness within different segments of the population but also the potential drivers behind this trend.
-Existing research suggests a complex interplay of factors, including economic considerations, rising educational attainment, changing social norms related to career and family life, and individual preferences.
+Existing research suggests a complex interplay of factors, including economic considerations, rising educational attainment, changing social norms related to career and family life, and individual preferences [@voluntaryChild].
 
-This paper aims to unpack these potential contributing factors and explore their impact on the rise of childlessness in America.
-By analyzing the data, I seek to answer crucial questions: What are the key factors contributing to this trend?
-How does childlessness manifest differently across various demographic groups?
-Finally, what are the potential social, economic, and individual consequences of this evolving family structure in the United States?
-Understanding these dynamics is crucial for shaping informed policy discussions and navigating the implications of childlessness in a rapidly changing world.
+This paper aims to unpack these factors and explore their impact on the rise of childlessness in America.
+By analyzing the data, we seek to answer crucial questions: What are the key factors contributing to this trend?
+How does childlessness manifest differently across demographic groups?
+Finally, what are the potential social, economic, and individual consequences of this new family structure in the United States?
+Understanding these dynamics is crucial in having an informed policy discussions and navigating the implications of childlessness in a rapidly changing world.
+
+Structurally, in @sec-intro, our paper starts with a brief dissection of the nature of childlessness in America and the contributing factors.
+Subsequently, in @sec-data and @sec-result, we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
+Next, @sec-discussion provides further insights and future areas of study.
+Finally, @sec-data concludes and summarizes our main findings.
+To complete the paper, @sec-appendix clarifies the variables within each dataset that we've used in our paper.
 
@@
 
-# Data
-This paper utilizes a multifaceted approach, drawing on various datasets to illuminate the complexities of childlessness in the United States.
+# Data {#sec-data}
+This paper uses a multifaceted approach, drawing on various datasets to understand the complexities of childlessness in the United States.
 One primary source is the General Social Survey (GSS), a long-running and highly respected survey conducted by the National Opinion Research Center (NORC) at the University of Chicago [@gss], from 1972 to 2022.
@@
 The 2022 iteration of the GSS, for the first time in its history, incorporated a combination of in-person, web self-administered, and phone interview methods.
-This expansion in data collection methods offers a potentially richer and more diverse sample, while acknowledging that specific variables might be limited to certain modes.
+This expansion in data collection methods offers a potentially richer and more diverse sample.
 The GSS team is transparent about these limitations and welcomes feedback from the research community to continuously improve their methodology.
 
-From the GSS, I utilize several key variables relevant to my analysis:
+From the GSS, we utilize several key variables relevant to our analysis:
 
@@
 * Age: This variable, categorized into relevant age groups (e.g., 18-34, 35-49, etc.), enables us to understand how childlessness manifests across different age demographics.
-It is important to note that the GSS data underwent specific recoding procedures for this analysis.
-For instance, the original "CHILD" variable was recoded to create the "childs" variable used here, and age categories were grouped for visualization purposes.
+It is important to note that the GSS data underwent specific naming procedures for this analysis.
+For instance, the original "CHILD" variable was renamed into the the "childs" variable used here, and age categories were grouped for visualization purposes.
 
@@
 ### Exploring Socioeconomic Class and Educational Attainment
-To gain a deeper understanding of the factors influencing these trends, I delve into additional data sets from NORC (@gss) examining the relationship between childlessness and socioeconomic class @tbl-nochild-class and educational attainment @tbl-nochild-degree.
+To gain a deeper understanding of the factors influencing these trends, we look into additional data sets from NORC (@gss) examining the relationship between childlessness and socioeconomic class @tbl-nochild-class and educational attainment @tbl-nochild-degree.
 
-From the GSS, I utilize several key variables to enrich my analysis and explore the potential influence of various factors on childbearing decisions:
+From the GSS, we utilize several key variables to enrich our analysis and explore the potential influence of various factors on childbearing decisions:
 
@@
 
-\newpage
 ```{r}
@@
 
+### Exploring Health Conditions and Marital Status
+
+In addition, additional data sets from NORC [@gss] were also leveraged to understand the context of childlessness in America from 1972-2022.
+@tbl-nochild-health and @tbl-nochild-marital present an interesting connection between childlessness and health conditions and martial status.
+
+From the GSS, we utilize several key variables relevant to our analysis:
+
+* Health Conditions: This variable, measured by medical check-ups of individuals who have been surveyed, will be used to clarify the the context behind childlessness of those with varying health status.
+* Marital Status: This variable enables us to see if marriage is a determining factor in childbearing decisions.
+
+@tbl-nochild-health uncovers the relationship between health conditions and childlessness.
+Intriguingly, those with 'excellent' and 'good' health are substantially more likely to be childless than those with 'fair' and 'poor' health.
+Several factors are attributable to this fact, but foremost careeer, lifestyle differences and financial considerations might come into play.
+Though, more data is needed to clarify the context of this trend, especially coupling it with variables like educational attainment or socioeconomic class.
+In addition, sampling bias might factor into this trend, where it's assumable that people with excellent health are often the younger population and those with less desirable health are older.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-health
+
+# Selected years to show better on the paper
+random_years <- c("Condition of health", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_health_subset <- table_nochild_health |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_health_subset, 
+      caption = "Childless Percentage of Population by Health Conditions and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
+@tbl-nochild-marital explores people's marital status and childlessness.
+Trends of childlessness vary with dataset.
+Unsurprisingly, people who are never married have the highest percentage of childlessness.
+However, those who are divorced, widowed and separated have fluctuating percentages of childlessness, which is understandable due to the complex emotional nature of these marital statuses.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-marital
+
+# Selected years to show better on the paper
+random_years <- c("Marital status", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_marital_subset <- table_nochild_marital |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_marital_subset, 
+      caption = "Childless Percentage of Population by Marital Status and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
 ## Data Limitations
@@
 
-# Result and Analysis
+### Decontextualized variables
 
-@fig-nochild-age illustrates of @tbl-nochild-age which is a consistent rise in childlessness across all age groups in the United States from 1972 to 2022.
+While these datasets aim to illuminate on the general time trends of childlessness with each variable measured, they are often de-contextualized for the same purpose.
+For instance, in @tbl-nochild-marital, there is no clarifying note on whether people who declared themselves as 'divorced' are divorced once or more.
+This can be extremely narrowing, and exclude nuances when analyzing the contexts behind childlessness.
+
+# Result and Analysis {#sec-result}
+
+@fig-nochild-age illustrates @tbl-nochild-age, which shows a consistent rise in childlessness across all age groups in the United States from 1972 to 2022.
 
@@
 People and households are divided into groups based on their socioeconomic status.
-The lower class experiences economic hardship, while the Working Class earns limited wages and faces job insecurity.
+The lower class experiences economic hardship, while the working class earns limited wages and faces job insecurity.
 The middle class has white-collar or professional occupations and enjoys a comfortable lifestyle.
@@
 
-While all classes show an increase in @fig-nochild-class, the data suggests the most significant rise occurred among the Lower Class, with their percentage of childless individuals nearly doubling from approximately 17.16% in 1972 to around 34.46% in 2022, marking a substantial increase of approximately 101.04% over the specified period.
-The Working Class also experienced a significant increase, from around 21.83% in 1972 to approximately 31.44% in 2022, indicating a rise of roughly 43.98%.
-Additionally, the Middle Class saw a slight increase, from approximately 28.24% in 1972 to about 30.09% in 2022, representing a rise of approximately 6.55%.
-Similarly, the Upper Class saw an increase, with their percentage of childless individuals climbing from approximately 20.24% in 1972 to about 27.87% in 2022, indicating an increase of roughly 37.71%.
+While all classes show an increase in @fig-nochild-class, the data suggests the most significant rise occurred among the lower class, with their percentage of childless individuals nearly doubling from approximately 17.16% in 1972 to around 34.46% in 2022, marking a substantial increase of approximately 101.04% over the specified period.
+The working class also experienced a significant increase, from around 21.83% in 1972 to approximately 31.44% in 2022, indicating a rise of roughly 43.98%.
+Additionally, the middle class saw a slight increase, from approximately 28.24% in 1972 to about 30.09% in 2022, representing a rise of approximately 6.55%.
+Similarly, the upper class saw an increase, with their percentage of childless individuals climbing from approximately 20.24% in 1972 to about 27.87% in 2022, indicating an increase of roughly 37.71%.
 
@@
 Individuals with a high school diploma consistently exhibit the highest percentage of childlessness throughout the entire period, ranging from around 28.15% in 1972 to 34.56% in 2022.
-Conversely, The percentage of childless individuals with a College and higher degree has decreased from 32.85% in 1972 to 29.84% in 2022.
+Conversely, The percentage of childless individuals with a college and higher degree has decreased from 32.85% in 1972 to 29.84% in 2022.
 The trend for individuals with a with less than a high school education increased from 16.97% in 1972 to 22.90% in 2022.
@@
 
-@fig-nochild-gender illustrates of @tbl-nochild-sex, explores gender differences in childlessness within the United States over a 50-year period (1972-2022).
+@fig-nochild-gender illustrates @tbl-nochild-sex, which explores gender differences in childlessness within the United States over a 50-year period (1972-2022).
 The graph utilizes line trends to depict the percentage of childless individuals categorized by gender (Male and Female).
@@
 
+@fig-nochild-health visualizes @tbl-nochild-health, which explores health differences in childlessness within the United States over a 50-year period (1972-2022).
+The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (Excellent, Good, Fair Poor).
+
+Overall, the graph illustrates an increase in childlessness accross all health levels and conditions approaching 2022.
+As mentioned, people with excellent health has the highest percentage of childlessness, having a 10.49% increase throughout the 50-year period (28.26% in 1972 to 38.75% in 2022).
+People with good and fair health exhibit similar rate of childlessness in 2022, at about 30%.
+In particular, people with fair health markedHowever, people with poor health, once having the lowest rate of childlessness (at 12.27% in 1972), now increases to 20.40% in 2022, which translates to a substantial 8.13% increase.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-health
+#| fig-cap: "Percentage Childless over Time by Health Conditions"
+
+# Pivot the data from wide to long format
+data_long <- pivot_longer(table_nochild_health, -`Condition of health`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long$Year <- as.integer(data_long$Year)
+
+# Plot the data
+ggplot(data_long, aes(x = Year, y = Percentage, color = `Condition of health`)) +
+  geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Health", color = "Condition of health") +
+  theme_minimal()
+
+```
+
+@fig-nochild-marital illustrates @tbl-nochild-marital, which explores marital status of people who are childless over a span of 50 years in the United States (1972-2022).
+The graph contains trend lines of varying marital statuses-Divorced, Married, Never married, Separated and Widowed-and how each has changed over the years in terms of childbearing status.
+
+The data reveals a gradual increase in childlessness for both genders over the past five decades.
+However, females consistently exhibit a lower percentage of childlessness compared to males throughout the entire period.
+While the gap appears to be slightly narrowing over time, females still have a lower percentage by around 8 percentage points in 2022 (approximately 28.32% for females and 36.09% for males).
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-marital
+#| fig-cap: "Percentage Childless over Time by Marital Status"
+
+# Pivot the data from wide to long format
+data_long <- pivot_longer(table_nochild_marital, -`Marital status`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long$Year <- as.integer(data_long$Year)
+
+# Plot the data
+ggplot(data_long, aes(x = Year, y = Percentage, color = `Marital status`)) +
+  geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Marital Status", color = "Marital status") +
+  theme_minimal()
+
+```
+
 The figures presented in this paper were generated using data from the [@gss].
@@
 It is important to remember that these figures solely present descriptive data and do not establish causal relationships.
-The Discussion section will delve deeper into potential explanations for the observed trends and explore the broader implications of these findings.
+The Discussion section will further analyze the observed trends and explore the broader implications of these findings.
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-# Conclusion
+# Conclusion {#sec-conclusion}
+
 This paper investigated the rising trend of childlessness in the United States, highlighting its prevalence across various demographics.
@@
 
-# Appendix
+# Appendix {#sec-appendix}
 The survey questions are available here under each trend: https://gssdataexplorer.norc.org/trends
@@
 
+(@) RESPONDENT'S MARITAL STATUS:
+- Married
+- Widowed
+- Divorced
+- Separated
+- Never married
+
+(@) RESPONDENT'S HEALTH CONDITION:
+- Excellent
+- Good
+- Fair
+- Poor
+
+\newpage
+
 # References

### 2024-03-11 scripts/03-test_data.R

@@ -12,6 +12,7 @@
 #### Workspace setup ####
 library(tidyverse)
 library(here)
+library(readr)
 
 #### Read in cleaned data sets ####
 table_nochild_age <-

### 2024-03-11 outputs/paper/paper.qmd
new formatting updates
new paragraph updates

@@
 Finally, @sec-data concludes and summarizes our main findings.
-To complete the paper, @sec-appendix clarifies the variables within each dataset that we've used in our paper.
+To complete the paper, @sec-appendix clarifies how each variable within each dataset is measured.
 
@@
 This trend suggests a shift in childbearing patterns, with individuals potentially delaying or forgoing parenthood altogether.
+
+\newpage
 
@@
 @tbl-nochild-marital explores people's marital status and childlessness.
-Trends of childlessness vary with dataset.
+Trends of childlessness vary with this dataset.
 Unsurprisingly, people who are never married have the highest percentage of childlessness.
@@
 
-@fig-nochild-gender illustrates @tbl-nochild-sex, which explores gender differences in childlessness within the United States over a 50-year period (1972-2022).
+@fig-nochild-gender explores gender differences in childlessness within the United States over a 50-year period (1972-2022).
 The graph utilizes line trends to depict the percentage of childless individuals categorized by gender (Male and Female).
@@
 
-@fig-nochild-health visualizes @tbl-nochild-health, which explores health differences in childlessness within the United States over a 50-year period (1972-2022).
+@fig-nochild-health explores health differences in childlessness within the United States over a 50-year period (1972-2022).
 The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (Excellent, Good, Fair Poor).
 
-Overall, the graph illustrates an increase in childlessness accross all health levels and conditions approaching 2022.
+Overall, the data illustrates an increase in childlessness accross all health levels and conditions approaching 2022.
 As mentioned, people with excellent health has the highest percentage of childlessness, having a 10.49% increase throughout the 50-year period (28.26% in 1972 to 38.75% in 2022).
@@
 
-@fig-nochild-marital illustrates @tbl-nochild-marital, which explores marital status of people who are childless over a span of 50 years in the United States (1972-2022).
+@fig-nochild-marital explores marital status of people who are childless over a span of 50 years in the United States (1972-2022).
 The graph contains trend lines of varying marital statuses-Divorced, Married, Never married, Separated and Widowed-and how each has changed over the years in terms of childbearing status.
 
-The data reveals a gradual increase in childlessness for both genders over the past five decades.
-However, females consistently exhibit a lower percentage of childlessness compared to males throughout the entire period.
-While the gap appears to be slightly narrowing over time, females still have a lower percentage by around 8 percentage points in 2022 (approximately 28.32% for females and 36.09% for males).
+The data demonstrates that a significant portion of people who have never married are childless.
+However, more and more people are reluctant to get married but still consider having children.
+This is observable in the childlessness percentage decrease of people who never marry, decreasing from 96.73% in 1972 to 72.53% in 2022.
+For people who are divorced, separated and widowed, this percentage flucatuates, displaying the complex dynamics between these relationship status and childbearing.
+For married people though, this childlessness percentage also decreases, from 10.51% in 1972 to 13.97% in 2022.
+This again demonstrates the overall decrease of childbirths due to varying factors.
 
@@
 
-(@) RESPONDENT'S MARITAL STATUS:
+(@) Are you currently--married, widowed, divorced, separated, or have you never been married?:
 - Married
@@
 
-(@) RESPONDENT'S HEALTH CONDITION:
+(@) Would you say your own health, in general, is excellent, good, fair, or  poor?:
 - Excellent

### 2024-03-11 outputs/paper/paper.qmd
new updates on formatting

@@
 
-Structurally, in @sec-intro, our paper starts with a brief dissection of the nature of childlessness in America and the contributing factors.
-Subsequently, in @sec-data and @sec-result, we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
-Next, @sec-discussion provides further insights and future areas of study.
-Finally, @sec-data concludes and summarizes our main findings.
-To complete the paper, @sec-appendix clarifies how each variable within each dataset is measured.
+Structurally, in [Introduction], our paper starts with a brief dissection of the nature of childlessness in America and the contributing factors.
+Subsequently, in [Data] and [Result], we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
+Next, [Discussion] provides further insights and future areas of study.
+Finally, [Conclusion] concludes and summarizes our main findings.
+To complete the paper, [Appendix] clarifies how each variable within each dataset is measured.
 
@@
 
-# Result and Analysis {#sec-result}
+# Result {#sec-result}
 
-@fig-nochild-age illustrates @tbl-nochild-age, which shows a consistent rise in childlessness across all age groups in the United States from 1972 to 2022.
+@fig-nochild-age shows a consistent rise in childlessness across all age groups in the United States from 1972 to 2022.
 
@@
 
-@fig-nochild-class visualizes @tbl-nochild-class which depicts the gradual increase in the percentage of childless individuals across four social classes (Lower Class, Working Class, Middle Class, Upper Class) in the United States from 1972 to 2022.
+@fig-nochild-class depicts the gradual increase in the percentage of childless individuals across four social classes (Lower Class, Working Class, Middle Class, Upper Class) in the United States from 1972 to 2022.
 The graph reveals an upward trend in childlessness for all classes over the five decades.
@@
 
-@fig-nochild-degree visualization of @tbl-nochild-degree, presents a comparative analysis of childlessness across individuals with different levels of educational attainment (Less than High School, High School and College's Degree or Higher) in the United States from 1972 to 2022.
+@fig-nochild-degree presents a comparative analysis of childlessness across individuals with different levels of educational attainment (Less than High School, High School and College's Degree or Higher) in the United States from 1972 to 2022.
 The graph utilizes line trends to depict the percentage of childless individuals within each degree category over the five decades.
@@
 
-Overall, the data illustrates an increase in childlessness accross all health levels and conditions approaching 2022.
+Overall, the data shows an increase in childlessness across all health levels and conditions approaching 2022.
 As mentioned, people with excellent health has the highest percentage of childlessness, having a 10.49% increase throughout the 50-year period (28.26% in 1972 to 38.75% in 2022).
@@
 
+\newpage
+
 @fig-nochild-marital explores marital status of people who are childless over a span of 50 years in the United States (1972-2022).
@@
 This is observable in the childlessness percentage decrease of people who never marry, decreasing from 96.73% in 1972 to 72.53% in 2022.
-For people who are divorced, separated and widowed, this percentage flucatuates, displaying the complex dynamics between these relationship status and childbearing.
+For people who are divorced, separated and widowed, this percentage fluctuates, displaying the complex dynamics between these relationship status and childbearing.
 For married people though, this childlessness percentage also decreases, from 10.51% in 1972 to 13.97% in 2022.
-This again demonstrates the overall decrease of childbirths due to varying factors.
 

### 2024-03-11 outputs/paper/paper.qmd
Added the names of the contributors

@@
 title: "Understanding the Landscape of Childlessness in the U.S.: A Data-Driven Exploration (1972-2022)"
-author: Mehrnoush Mohammadi & Quang Mai
+author: 
+  - Mehrnoush Mohammadi 
+  - Quang Mai
 thanks: "Code and data are available at: https://github.com/mehrnoush68/childlessness-in-america.git"
@@
 
-(@) Are you currently--married, widowed, divorced, separated, or have you never been married?:
+(@) Are you currently:
 - Married
@@
 
-(@) Would you say your own health, in general, is excellent, good, fair, or  poor?:
+(@) Would you say your own health, in general:
 - Excellent

### 2024-03-11 scripts/00-simulate_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Simulates data for the data sets no child, age, class identification, degree, and gender
-# Author: Mehrnoush Mohammadi
+# Author: Mehrnoush Mohammadi & Quang Mai
 # Date: 2 March 2024
 # Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT

### 2024-03-11 scripts/01-download_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Downloads the data from US General Social Survey
-# Author: Mehrnoush Mohammadi
+# Author: Mehrnoush Mohammadi & Quang Mai
 # Date: 2 March 2024
 # Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT

### 2024-03-11 scripts/02-data_cleaning.R

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Cleans the raw plane data childlessness in America
-# Author: Mehrnoush Mohammadi
+# Author: Mehrnoush Mohammadi & Quang Mai
 # Date: 2 March 2024
 # Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT

### 2024-03-11 scripts/03-test_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Tests the three data sets
-# Author: Mehrnoush Mohammadi
+# Author: Mehrnoush Mohammadi & Quang Mai
 # Date: 2 March 2024
 # Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT

### 2024-03-11 outputs/paper/paper.qmd
Fig's labs font size changed and some Typo fixed

@@
 This paper aims to unpack these factors and explore their impact on the rise of childlessness in America.
-By analyzing the data, we seek to answer crucial questions: What are the key factors contributing to this trend?
+By analyzing the data, we seek to answer crucial questions: What are some important factors contributing to this trend?
 How does childlessness manifest differently across demographic groups?
 Finally, what are the potential social, economic, and individual consequences of this new family structure in the United States?
-Understanding these dynamics is crucial in having an informed policy discussions and navigating the implications of childlessness in a rapidly changing world.
+Understanding these dynamics is crucial in having an informed policy discussion and navigating the implications of childlessness in a rapidly changing world.
 
@@
 * Children: This variable captures the number of children ever born alive to the respondent, providing a direct measure of childlessness (having no children).
-* Year: This variable allows to analyze trends in childlessness over time, encompassing the chosen time frame for the survey.
+* Year: This variable allows us to analyze trends in childlessness over time, encompassing the chosen time frame for the survey.
 * Age: This variable, categorized into relevant age groups (e.g., 18-34, 35-49, etc.), enables us to understand how childlessness manifests across different age demographics.
-It is important to note that the GSS data underwent specific naming procedures for this analysis.
-For instance, the original "CHILD" variable was renamed into the the "childs" variable used here, and age categories were grouped for visualization purposes.
 
@@
 * Socioeconomic Class: Data on socioeconomic class, often measured by income or occupation, will be utilized to understand the impact of economic factors on childbearing choices.
-* Educational Attainment: Data on educational attainment, such as highest degree obtained, will be incorporated to explore the relationship between educational level and childbearing.
-This data can be obtained from sources like the National Center for Education Statistics.
+* Educational Attainment: Data on educational attainment, such as the highest degree obtained, will be incorporated to explore the relationship between educational level and childbearing.
 
 @tbl-nochild-class reveals a potential link between socioeconomic class and childlessness.
-The data suggests that individuals in the lower class and working class exhibit higher percentages of childlessness compared to those in the middle class and upper class.
+The data suggests that individuals in the lower and working classes exhibit higher percentages of childlessness compared to those in the middle and upper classes.
 This pattern might be attributed to various factors, such as financial constraints, job insecurity, or limited access to childcare resources experienced by individuals in lower socioeconomic classes.
@@
 
-From the GSS, we utilize several key variables relevant to our analysis:
+From the GSS, we utilize several important variables relevant to our analysis:
 
-* Health Conditions: This variable, measured by medical check-ups of individuals who have been surveyed, will be used to clarify the the context behind childlessness of those with varying health status.
+* Health Conditions: This variable, measured by medical check-ups of individuals who have been surveyed, will be used to clarify the context behind the childlessness of those with varying health statuses.
 * Marital Status: This variable enables us to see if marriage is a determining factor in childbearing decisions.
@@
 Intriguingly, those with 'excellent' and 'good' health are substantially more likely to be childless than those with 'fair' and 'poor' health.
-Several factors are attributable to this fact, but foremost careeer, lifestyle differences and financial considerations might come into play.
-Though, more data is needed to clarify the context of this trend, especially coupling it with variables like educational attainment or socioeconomic class.
+Several factors are attributable to this fact, but foremost, career, lifestyle differences and financial considerations might come into play.
+However, more data is needed to clarify the context of this trend, especially coupling it with variables like educational attainment or socioeconomic class.
 In addition, sampling bias might factor into this trend, where it's assumable that people with excellent health are often the younger population and those with less desirable health are older.
@@
   geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "solid", size = 1.2) +  # Add running average trend line
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
   labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Age Group", color = "Age Group") +
-  theme_minimal()
+  theme_minimal() +
+  theme(text = element_text(size = 10))
 
@@
 It is important to acknowledge, however, that the data series has gaps for several years within this timeframe (1979, 1981, 1992, 1995, 1997, 1999, 2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017, and 2019).
-These missing data points can potentially limit the generalization of the observed trends and hinder a comprehensive understanding of the childlessness over time.
+These missing data points can potentially limit the generalization of the observed trends and hinder a comprehensive understanding of childlessness over time.
 Despite these limitations, the available data reveals a steady increase in the percentage of individuals without children across all age groups.
@@
 
-@fig-nochild-class depicts the gradual increase in the percentage of childless individuals across four social classes (Lower Class, Working Class, Middle Class, Upper Class) in the United States from 1972 to 2022.
+@fig-nochild-class depicts the gradual increase in the percentage of childless individuals across four social classes (lower class, working class, middle class, and upper class) in the United States from 1972 to 2022.
 The graph reveals an upward trend in childlessness for all classes over the five decades.
@@
   geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "solid", size = 1.2) +  # Add running average trend line
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
   labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Class", color = "Breakdown Category") +
-  theme_minimal()
+  theme_minimal() +
+  theme(text = element_text(size = 10))
 
@@
 
-@fig-nochild-degree presents a comparative analysis of childlessness across individuals with different levels of educational attainment (Less than High School, High School and College's Degree or Higher) in the United States from 1972 to 2022.
+@fig-nochild-degree presents a comparative analysis of childlessness across individuals with different levels of educational attainment (less than high school, high school and college degree or higher) in the United States from 1972 to 2022.
 The graph utilizes line trends to depict the percentage of childless individuals within each degree category over the five decades.
@@
 Individuals with a high school diploma consistently exhibit the highest percentage of childlessness throughout the entire period, ranging from around 28.15% in 1972 to 34.56% in 2022.
-Conversely, The percentage of childless individuals with a college and higher degree has decreased from 32.85% in 1972 to 29.84% in 2022.
+Conversely, the percentage of childless individuals with a college and higher degree has decreased from 32.85% in 1972 to 29.84% in 2022.
 The trend for individuals with a with less than a high school education increased from 16.97% in 1972 to 22.90% in 2022.
@@
   labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Degree", color = "Degree") +
-  theme_minimal()
+  theme_minimal() +
+  theme(text = element_text(size = 10))
 
@@
   labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Gender", color = "Gender") +
-  theme_minimal()
+  theme_minimal() +
+  theme(text = element_text(size = 10))
 
@@
 @fig-nochild-health explores health differences in childlessness within the United States over a 50-year period (1972-2022).
-The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (Excellent, Good, Fair Poor).
+The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (excellent, good, fair, and poor).
 
 Overall, the data shows an increase in childlessness across all health levels and conditions approaching 2022.
-As mentioned, people with excellent health has the highest percentage of childlessness, having a 10.49% increase throughout the 50-year period (28.26% in 1972 to 38.75% in 2022).
-People with good and fair health exhibit similar rate of childlessness in 2022, at about 30%.
-In particular, people with fair health markedHowever, people with poor health, once having the lowest rate of childlessness (at 12.27% in 1972), now increases to 20.40% in 2022, which translates to a substantial 8.13% increase.
-
+As mentioned, people with excellent health have the highest percentage of childlessness, having a 10.49% increase throughout the 50-year period (28.26% in 1972 to 38.75% in 2022).
+People with good and fair health exhibit a similar rate of childlessness in 2022, at about 30%.
+In particular, people with fair health marked; however, people with poor health, once having the lowest rate of childlessness (at 12.27% in 1972), now increases to 20.40% in 2022, which translates to a substantial 8.13% increase.
 ```{r}
@@
   labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Health", color = "Condition of health") +
-  theme_minimal()
+  theme_minimal() +
+  theme(text = element_text(size = 10))
 
@@
 
-@fig-nochild-marital explores marital status of people who are childless over a span of 50 years in the United States (1972-2022).
+@fig-nochild-marital explores the marital status of people who are childless over a span of 50 years in the United States (1972-2022).
 The graph contains trend lines of varying marital statuses-Divorced, Married, Never married, Separated and Widowed-and how each has changed over the years in terms of childbearing status.
@@
 This is observable in the childlessness percentage decrease of people who never marry, decreasing from 96.73% in 1972 to 72.53% in 2022.
-For people who are divorced, separated and widowed, this percentage fluctuates, displaying the complex dynamics between these relationship status and childbearing.
+For people who are divorced, separated and widowed, this percentage fluctuates, displaying the complex dynamics between these relationship statuses and childbearing.
 For married people though, this childlessness percentage also decreases, from 10.51% in 1972 to 13.97% in 2022.
@@
   labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Marital Status", color = "Marital status") +
-  theme_minimal()
+  theme_minimal() +
+  theme(text = element_text(size = 10))
 
@@
 The figures presented in this paper were generated using data from the [@gss].
-These figures visually represent the key findings related to the trends and patterns in childlessness across different age, social classes, educational attainment levels, and genders in the United States over the past five decades (1972-2022).
+These figures visually represent the key findings related to the trends and patterns in childlessness across different ages, social classes, educational attainment levels, and genders in the United States over the past five decades (1972-2022).
 It is important to remember that these figures solely present descriptive data and do not establish causal relationships.
@@
 
-To gain a deeper understanding of childlessness, there are some crucial next steps that need to be taken.
+To gain a deeper understanding of childlessness, some crucial next steps need to be taken.
 Firstly, employing longitudinal studies will allow the exploration of causal relationships between factors and childbearing decisions over time.

### 2024-03-12 outputs/paper/paper.qmd
Survey added

@@
 # Appendix {#sec-appendix}
-The survey questions are available here under each trend: https://gssdataexplorer.norc.org/trends
+The supplementary survey is accessible through this link: https://forms.gle/DsFgJCKPeCtdqZKL6
+
+## Survey Preamble
+The General Social Survey (GSS) collects data on social trends to help track changes in Canadians' living conditions and well-being over time.
+Also, it provides information on specific social policy issues that are currently or potentially important.
+
+We are conducting a study to investigate the reasons behind childlessness in different groups of people.
+To do so, we will analyze various data sets, including information on age, gender, education, socioeconomic status, health conditions, and marital status.
+By examining these factors, we hope to gain insight into how they impact rates of childbearing.
+
+Participation in this survey is voluntary, and all responses will be kept anonymous.
+If you choose to participate, you will have the freedom to skip any questions and can withdraw from the survey at any time.
+Your individual responses will be recorded in our data set with a unique identification number for analysis purposes only.
+
+Thank you for taking the time to complete this survey. If you have any questions or require further information, please feel free to contact mm.mohammadi@mail.utoronto.ca.
 
 ## Survey Questions from GSS
+
+(@) How old are you?
+(short answer text)
+
+(@) Gender:
+- Female
+- Male
+- Prefer not to say
+
+(@) Are you currently:
+- Married
+- Widowed
+- Divorced
+- Separated
+- Never married
+
+(@) What is your highest degree?
+- College +
+- High school
+- Less than high school
+
 (@) How many children have you ever had?  Please count all that were born alive at any time (including any you had from a previous marriage).
@@
 
-(@) RESPONDENT'S AGE
-(short answer text)
-
 (@) If you were asked to use one of four names for your social class, which would you say you belong in: 
@@
 - The upper class
-
-(@) RESPONDENT'S DEGREE:
-- College +
-- High school
-- Less than high school
-
-(@) RESPONDENT'S SEX:
-- Female
-- Male
-
-(@) Are you currently:
-- Married
-- Widowed
-- Divorced
-- Separated
-- Never married
 

### 2024-03-13 scripts/00-simulate_data.R
Update on the simulate data

@@ -10,65 +10,34 @@
 library(tidyverse)
 
 #### Simulate data ####
-# Simulated data for table_nochild_age
+
 set.seed(123)  # For reproducibility
-age_groups <- c("18-34", "35-49", "50-64", "65+")
-years <- 1972:2022
 
-# Generate random percentages for each age group and year
-percentages_age <- matrix(runif(length(age_groups) * length(years), min = 5, max = 70), nrow = length(age_groups))
-colnames(percentages_age) <- years
-rownames(percentages_age) <- age_groups
+# Function to generate simulated data
+generate_data <- function(labels, years, n) {
+  tibble(
+    Category = rep(labels, each = length(years)),
+    Year = rep(years, length(labels)),
+    Percentage = runif(length(labels) * length(years), min = 5, max = 70),
+    Table = n
+  )
+}
 
-print(percentages_age)
+# Create a tibble of label vectors for each table
+labels <- list(
+  age = c("18-34", "35-49", "50-64", "65+"),
+  class = c("Lower class", "Working class", "Middle class", "Upper class"),
+  degree = c("High school diploma", "Bachelor's degree", "Master's degree", "Doctoral degree"),
+  sex = c("Male", "Female"),
+  health = c("Excellent", "Good", "Fair", "Poor"),
+  marital = c("Never married", "Married", "Separated", "Divorced", "Widowed")
+)
 
-# Simulated data for table_nochild_class
-class_labels <- c("Lower class", "Working class", "Middle class", "Upper class")
+# Generate simulated data for each table using map_dfr
+simulated_data <- imap_dfr(labels, ~generate_data(.x, 1972:2022, .y))
 
-# Generate random percentages for each class and year
-percentages_class <- matrix(runif(length(class_labels) * length(years), min = 5, max = 70), nrow = length(class_labels))
-colnames(percentages_class) <- years
-rownames(percentages_class) <- class_labels
+# Print the simulated data for each table
+#print(simulated_data, n = 1173) this is the whole simulated data
+print(simulated_data)
 
-print(percentages_class)
-
-# Simulated data for table_nochild_degree
-degree_labels <- c("High school diploma", "Bachelor's degree", "Master's degree", "Doctoral degree")
-
-# Generate random percentages for each degree and year
-percentages_degree <- matrix(runif(length(degree_labels) * length(years), min = 5, max = 70), nrow = length(degree_labels))
-colnames(percentages_degree) <- years
-rownames(percentages_degree) <- degree_labels
[…]

### 2024-03-13 scripts/01-download_data.R
Download data info added

@@ -7,42 +7,10 @@
 # Pre-requisites: Download the raw data from US General Social Survey (https://gssdataexplorer.norc.org/trends)
 
 
-#### Workspace setup ####
-library(tidyverse)
-library(readr)
-
-# Note: data is behind sign-in
-# data can be accessed at https://gssdataexplorer.norc.org/trends from the US General Social Survey
-# data was downloaded as .xlsx file format and saved as .csv format on the computer
-
-#### Load data ####
-nochild_age_data <-
-  read_csv(
-    file = "inputs/data/nochild_age.csv",
-    show_col_types = FALSE,
-  )
-nochild_class_data <-
-  read_csv(
-    file = "inputs/data/nochild_class.csv",
-    show_col_types = FALSE,
-  )
-nochild_sex_data <-
-  read_csv(
-    file = "inputs/data/nochild_sex.csv",
-    show_col_types = FALSE,
-  )
-nochild_degree_data <-
-  read_csv(
-    file = "inputs/data/nochild_higestdegree.csv",
-    show_col_types = FALSE,
-  )
-nochild_health_data <-
-  read_csv(
-    file = "inputs/data/nochild_health.csv",
-    show_col_types = FALSE,
-  )
-nochild_marital_data <-
-  read_csv(
-    file = "inputs/data/nochild_marital.csv",
-    show_col_types = FALSE,
-  )
+# We accessed data from the US General Social Survey on Gender and Marriage trends at https://gssdataexplorer.norc.org/trends. 
+# The data is behind a sign-in.
+# To find the data, we went to the Gender and Marriage trends section and clicked on Children & Working link.
+# Then, we found different trends related to Children & Working, such as Gender & Marriage, Current Affairs, Civil Liberties, Politics, and Religion & Spirituality.
+# Our focus was on the Gender and Marriage section for individuals with no children, with different ages, degrees, conditions of health, sex, and marital status breakdown.
+# To download the data, we clicked on the export button and download the data in Excel format (.xlsx). 
+# We then saved it as a .csv format using Microsoft Excel.

### 2024-03-13 scripts/02-data_cleaning.R
Data cleaning file update

@@ -9,124 +9,37 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(readr)
-library(dplyr)
-library(knitr)
 
-#### Load data ####
-nochild_age_data <-
-  read_csv(
-    file = "inputs/data/nochild_age.csv",
-    show_col_types = FALSE,
-  )
-nochild_class_data <-
-  read_csv(
-    file = "inputs/data/nochild_class.csv",
-    show_col_types = FALSE,
-  )
-nochild_sex_data <-
-  read_csv(
-    file = "inputs/data/nochild_sex.csv",
-    show_col_types = FALSE,
-  )
-nochild_degree_data <-
-  read_csv(
-    file = "inputs/data/nochild_higestdegree.csv",
-    show_col_types = FALSE,
-  )
-nochild_health_data <-
-  read_csv(
-    file = "inputs/data/nochild_health.csv",
-    show_col_types = FALSE,
-  )
-nochild_marital_data <-
-  read_csv(
-    file = "inputs/data/nochild_marital.csv",
-    show_col_types = FALSE,
-  )
-
-# Cleaning Data
-data_cleaned_age <- nochild_age_data
-data_cleaned_class <- nochild_class_data
-data_cleaned_sex <- nochild_sex_data
-data_cleaned_degree <- nochild_degree_data
-data_cleaned_health <- nochild_health_data
-data_cleaned_marital <- nochild_marital_data
-
-
-
-# Define a function to clean the data
-clean_data_age <- function(x) {
-  as.numeric(gsub("\\s*\\(.*\\)", "", x))
-}
-clean_data_class <- function(x) {
-  as.numeric(gsub("\\s*\\(.*\\)", "", x))
-}
-clean_data_sex <- function(x) {
-  as.numeric(gsub("\\s*\\(.*\\)", "", x))
-}
-clean_data_degree <- function(x) {
-  as.numeric(gsub("\\s*\\(.*\\)", "", x))
-}
-clean_data_health <- function(x) {
-  as.numeric(gsub("\\s*\\(.*\\)", "", x))
-}
-clean_data_marital <- function(x) {
-  as.numeric(gsub("\\s*\\(.*\\)", "", x))
-}
-
-# Clean the data
-clean_data_age <- data_cleaned_age |>
-  mutate(across(-1, clean_data_age))
-
-# Rename the 'age' column to 'Age'
-clean_data_age <- clean_data_age |>
-  rename(Age = age)
-
-clean_data_class <- data_cleaned_class |>
[…]

### 2024-03-13 scripts/02-data_cleaning.R
Data cleaning file update - the code is shorter

@@ -9,6 +9,9 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(magrittr)
+library(readr)
+library(dplyr)
 
 # List of input files
 input_files <- c(
@@ -20,26 +23,27 @@ input_files <- c(
   "nochild_marital.csv"
 )
 
-# Remove "_nochild" from file names
-input_files <- gsub("_nochild", "", input_files)
+# Remove "nochild_" from input file names and create output file names
+output_files <- gsub("nochild_", "", input_files)
+
+# Remove "higest" from input file names and create output file names
+output_files[input_files == "nochild_higestdegree.csv"] <- "degree.csv"
 
 # List of output files
-output_files <- paste0("clean_data_", gsub("_nochild", "", tools::file_path_sans_ext(input_files)), ".csv")
-
+output_files <- paste0("clean_data_", tools::file_path_sans_ext(output_files), ".csv")
 
 
 # Loop over input files
-for (i in seq_along(input_files)) {
+for (input_file in input_files) {
   # Read data with show_col_types = FALSE
-  data <- read_csv(file.path("inputs/data", input_files[i]), show_col_types = FALSE)
+  data <- read_csv(file.path("inputs/data", input_file), show_col_types = FALSE)
   
   # Clean data
-  cleaned_data <- data %>%
+  cleaned_data <- data |>
     mutate(across(-1, ~as.numeric(gsub("\\s*\\(.*\\)", "", .))))
   
   # Write cleaned data to output file
-  output_file <- gsub("_nochild", "", output_files[i])
+  output_file <- output_files[input_files == input_file]
   write_csv(cleaned_data, file.path("outputs/data", output_file))
 }
-
 

### 2024-03-14 scripts/02-data_cleaning.R
Debugging the data cleaning file (Renaming age to Age)

@@ -33,6 +33,7 @@ output_files[input_files == "nochild_higestdegree.csv"] <- "degree.csv"
 output_files <- paste0("clean_data_", tools::file_path_sans_ext(output_files), ".csv")
 
 
+
 # Loop over input files
 for (input_file in input_files) {
   # Read data with show_col_types = FALSE
@@ -41,7 +42,12 @@ for (input_file in input_files) {
   # Clean data
   cleaned_data <- data |>
     mutate(across(-1, ~as.numeric(gsub("\\s*\\(.*\\)", "", .))))
-  
+
+  # Rename 'age' to 'Age' if 'age' exists in the column names
+  if (any(colnames(cleaned_data) == "age")) {
+    cleaned_data <- cleaned_data |>
+      rename(Age = age)
+  } 
   # Write cleaned data to output file
   output_file <- output_files[input_files == input_file]
   write_csv(cleaned_data, file.path("outputs/data", output_file))

### 2024-03-14 outputs/paper/paper.qmd
Updates on abstract, and survey questions

@@
 date-format: long
-abstract: "This paper investigates trends in childlessness in the United States from 1972 to 2022.
-Data from the General Social Survey (GSS) is analyzed to consider factors like age, gender, education, class, health condition, marital status and number of children.
-The analysis reveals a consistent rise in childlessness across various demographics, potentially linked to economic considerations, educational attainment, and shifting social norms.
-The paper explores the potential social and economic implications of this trend, highlighting the need for further research to navigate the complexities surrounding childlessness in today's society."
+abstract: "This paper investigates U.S.
+childlessness trends (1972-2022) using data collected from the General Social Survey (GSS).
+The research covers several factors such as age, gender, education, social class, health conditions, marital status, and number of children.
+The analysis reveals rising childlessness across demographics, potentially linked to economic factors, education, and social norms.
+We explore the social and economic implications of this trend and emphasize the need for further research on childlessness in modern America."
 format: pdf
@@
 The American family landscape is undergoing a significant transformation, marked by a growing trend of childlessness.
-Traditionally viewed as a societal norm, having children now becomes a choice.
+Traditionally viewed as a societal norm, having children now becomes a choice [@voluntaryChild].
 This paper looks into this phenomenon by examining various datasets on factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they affect childbearing rates over the past 50 years.
@@
 
-## Survey Questions from GSS
+## Survey Questions for Childlessness
 
-(@) How old are you?
+**Demographics**
+
+(@) Age:
 (short answer text)
@@
 - Male
+- Non-binary
 - Prefer not to say
 
-(@) Are you currently:
+(@) Marital Status:
 - Married
@@
 - Never married
+- Living in a committed partnership
 
-(@) What is your highest degree?
-- College +
-- High school
-- Less than high school
+(@) Highest Level of Education:
+- No formal education
+- Less than high school diploma
+- High school diploma
+- Some college degree
+- Associate's degree
+- Bachelor's degree
+- Master's degree
+- Doctorate or professional degree
 
-(@) How many children have you ever had?  Please count all that were born alive at any time (including any you had from a previous marriage).
-- 0
-- 1
-- 2
-- 3
-- 4
-- 5
-- 6
-- 7
-- 8 or more
+**Family and Childbearing**
 
-(@) If you were asked to use one of four names for your social class, which would you say you belong in: 
+(@) Do you currently have any children?
+- Yes
+- No
+
+(@) If yes, How many children have you ever had? Please count all that were born alive at any time (including any you had from a previous marriage).
+(short answer text)
+
+(@) At what age did you have your first child (if any)?
+(Short answer text)
+
+(@) Do you currently desire to have children? 
+- Yes, I desire to have children.
+- No, I do not desire to have children.
+- I already have children and would like more.
+- I have not decided about having children.
+
+(@) If you do not have children, what are the main reasons for your decision?
+- Financial concerns
+- Career aspirations
+- Not interested in parenthood
+- Health concerns
+- Other (short answer text)
+
+**Socioeconomic**
+
+(@) What is your current employment status?
+- Employed full-time
+- Employed part-time
+- Self-employed
+- Unemployed
+- Retired
+- Student
+- Other (short answer text)
+
+(@) Do you feel your current financial situation allows you to comfortably raise children?
+- Yes
+- No
+- I do not know
+
+(@) If you were asked to use one of four names for your social class, which would you say you belong in? 
 - The lower class 
@@
 
+(@) Do you or your partner have any significant health conditions that may impact your decision to have children?
+- Yes
+- No
+- Maybe
+
+**Thank you for your participation**
+
+We sincerely appreciate your time and thoughtful responses to our survey about childbearing decisions. If you have any additional comments or suggestions, please feel free to contact us.
+
 \newpage

### 2024-03-14 outputs/paper/paper.qmd
New Average graphs added for the result section

@@
 #### Load data ####
-table_nochild_age <-
-  read_csv(
-    file = "../data/clean_data_age.csv",
-    show_col_types = FALSE
-  )
+table_nochild_age <- read_csv(file = "../data/clean_data_age.csv",show_col_types = FALSE)
 
-table_nochild_class <-
-  read_csv(
-    file = "../data/clean_data_class.csv",
-    show_col_types = FALSE
-  )
+table_nochild_class <- read_csv(file = "../data/clean_data_class.csv",show_col_types = FALSE)
 
-table_nochild_degree <-
-  read_csv(
-    file = "../data/clean_data_degree.csv",
-    show_col_types = FALSE
-  )
+table_nochild_degree <- read_csv(file = "../data/clean_data_degree.csv",show_col_types = FALSE)
 
-table_nochild_sex <-
-  read_csv(
-    file = "../data/clean_data_sex.csv",
-    show_col_types = FALSE
-  )
+table_nochild_sex <- read_csv(file = "../data/clean_data_sex.csv",show_col_types = FALSE)
 
-table_nochild_marital <-
-  read_csv(
-    file = "../data/clean_data_marital.csv",
-    show_col_types = FALSE
-  )
+table_nochild_marital <- read_csv(file = "../data/clean_data_marital.csv",show_col_types = FALSE)
 
-table_nochild_health <-
-  read_csv(
-    file = "../data/clean_data_health.csv",
-    show_col_types = FALSE
-  )
+table_nochild_health <- read_csv(file = "../data/clean_data_health.csv",show_col_types = FALSE)
 
@@
 
-nochild_age_data_long <- tidyr::pivot_longer(table_nochild_age, -Age, names_to = "year", values_to = "percentage")
+data_long_age <- tidyr::pivot_longer(table_nochild_age, -Age, names_to = "year", values_to = "percentage")
 
-ggplot(nochild_age_data_long, aes(x = as.integer(year), y = percentage, color = Age)) +
+ggplot(data_long_age, aes(x = as.integer(year), y = percentage, color = Age)) +
   geom_line() +
@@
 # Pivot the data from wide to long format
-data_long <- pivot_longer(table_nochild_class, -`Class Identification`, names_to = "Year", values_to = "Percentage")
+data_long_class <- pivot_longer(table_nochild_class, -`Class Identification`, names_to = "Year", values_to = "Percentage")
 
 # Convert Year column to integer
-data_long$Year <- as.integer(data_long$Year)
+data_long_class$Year <- as.integer(data_long_class$Year)
 
 # Plot the data
-ggplot(data_long, aes(x = Year, y = Percentage, color = `Class Identification`)) +
+ggplot(data_long_class, aes(x = Year, y = Percentage, color = `Class Identification`)) +
   geom_line() +
@@
 # Pivot the data from wide to long format
-data_long <- pivot_longer(table_nochild_degree, -`Degree`, names_to = "Year", values_to = "Percentage")
+data_long_degree <- pivot_longer(table_nochild_degree, -`Degree`, names_to = "Year", values_to = "Percentage")
 
 # Convert Year column to integer
-data_long$Year <- as.integer(data_long$Year)
+data_long_degree$Year <- as.integer(data_long_degree$Year)
 
 # Plot the data
-ggplot(data_long, aes(x = Year, y = Percentage, color = `Degree`)) +
+ggplot(data_long_degree, aes(x = Year, y = Percentage, color = `Degree`)) +
   geom_line() +
@@
 # Pivot the data from wide to long format
-data_long <- pivot_longer(table_nochild_sex, -`Gender`, names_to = "Year", values_to = "Percentage")
+data_long_sex <- pivot_longer(table_nochild_sex, -`Gender`, names_to = "Year", values_to = "Percentage")
 
 # Convert Year column to integer
-data_long$Year <- as.integer(data_long$Year)
+data_long_sex$Year <- as.integer(data_long_sex$Year)
 
 # Plot the data
-ggplot(data_long, aes(x = Year, y = Percentage, color = `Gender`)) +
+ggplot(data_long_sex, aes(x = Year, y = Percentage, color = `Gender`)) +
   geom_line() +
@@
 # Pivot the data from wide to long format
-data_long <- pivot_longer(table_nochild_health, -`Condition of health`, names_to = "Year", values_to = "Percentage")
+data_long_health <- pivot_longer(table_nochild_health, -`Condition of health`, names_to = "Year", values_to = "Percentage")
 
 # Convert Year column to integer
-data_long$Year <- as.integer(data_long$Year)
+data_long_health$Year <- as.integer(data_long_health$Year)
 
 # Plot the data
-ggplot(data_long, aes(x = Year, y = Percentage, color = `Condition of health`)) +
+ggplot(data_long_health, aes(x = Year, y = Percentage, color = `Condition of health`)) +
   geom_line() +
@@
 # Pivot the data from wide to long format
-data_long <- pivot_longer(table_nochild_marital, -`Marital status`, names_to = "Year", values_to = "Percentage")
+data_long_marital <- pivot_longer(table_nochild_marital, -`Marital status`, names_to = "Year", values_to = "Percentage")
 
 # Convert Year column to integer
-data_long$Year <- as.integer(data_long$Year)
+data_long_marital$Year <- as.integer(data_long_marital$Year)
 
 # Plot the data
-ggplot(data_long, aes(x = Year, y = Percentage, color = `Marital status`)) +
+ggplot(data_long_marital, aes(x = Year, y = Percentage, color = `Marital status`)) +
   geom_line() +
@@
 The Discussion section will further analyze the observed trends and explore the broader implications of these findings.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-avg-marital
+#| fig-cap: "Average Childless Percentage of Population over Time by Marital Status"
+
+# Group the data by Year and calculate the average of Percentage for each year
+avg_data <- data_long_marital |>
+  group_by(Year) |>
+  summarise(Sum_Percentage = sum(Percentage, na.rm = TRUE)) |>
+  mutate(avg_Percentage = Sum_Percentage / 5)
+
+# Print the average data
+# print(avg_data)
+
+# Plot the data
+ggplot(avg_data, aes(x = Year, y = avg_Percentage)) +
+  geom_line() +
+  geom_point() +
+  labs(x = "Year", y = "Average Percentage of Population", title = "Average Childless Percentage of Population over Time by Marital Status") +
+  theme_minimal()
+
+```
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-avg-class
+#| fig-cap: "Average Percentage Childless over Time by Class"
+
+# Group the data by Year and calculate the average of Percentage for each year
+avg_data <- data_long_class |>
+  group_by(Year) |>
+  summarise(Sum_Percentage = sum(Percentage, na.rm = TRUE)) |>
+  mutate(avg_Percentage = Sum_Percentage / 4)
+
+# Print the average data
+# print(avg_data)
+
+# Plot the data
+ggplot(avg_data, aes(x = Year, y = avg_Percentage)) +
+  geom_line() +
+  geom_point() +
+  labs(x = "Year", y = "Average Percentage of Population", title = "Average Percentage Childless over Time by Class") +
+  theme_minimal()
+
+```
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-avg-gender
+#| fig-cap: "Average Percentage Childless over Time by Gender"
+
+# Group the data by Year and calculate the average of Percentage for each year
+avg_data <- data_long_sex |>
+  group_by(Year) |>
+  summarise(Sum_Percentage = sum(Percentage, na.rm = TRUE)) |>
+  mutate(avg_Percentage = Sum_Percentage / 2)
+
+# Print the average data
+# print(avg_data)
+
+# Plot the data
+ggplot(avg_data, aes(x = Year, y = avg_Percentage)) +
+  geom_line() +
+  geom_point() +
+  labs(x = "Year", y = "Average Percentage of Population", title = "Average Percentage Childless over Time by Gender") +
+  theme_minimal()
+
+```
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-avg-degree
+#| fig-cap: "Average Percentage Childless over Time by Degree"
+
+# Group the data by Year and calculate the average of Percentage for each year
+avg_data <- data_long_degree |>
+  group_by(Year) |>
+  summarise(Sum_Percentage = sum(Percentage, na.rm = TRUE)) |>
+  mutate(avg_Percentage = Sum_Percentage / 3)
+
+# Print the average data
+# print(avg_data)
+
+# Plot the data
+ggplot(avg_data, aes(x = Year, y = avg_Percentage)) +
+  geom_line() +
+  geom_point() +
+  labs(x = "Year", y = "Average Percentage of Population", title = "Average Percentage Childless over Time by Degree") +
+  theme_minimal()
+
+```
 

### 2024-03-14 scripts/00-simulate_data.R
Updated the dates

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulates data for the data sets no child, age, class identification, degree, and gender
 # Author: Mehrnoush Mohammadi & Quang Mai
-# Date: 2 March 2024
+# Date: 14 March 2024
 # Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT
 

### 2024-03-14 scripts/01-download_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads the data from US General Social Survey
 # Author: Mehrnoush Mohammadi & Quang Mai
-# Date: 2 March 2024
+# Date: 14 March 2024
 # Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Download the raw data from US General Social Survey (https://gssdataexplorer.norc.org/trends)

### 2024-03-14 scripts/02-data_cleaning.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw plane data childlessness in America
 # Author: Mehrnoush Mohammadi & Quang Mai
-# Date: 2 March 2024
+# Date: 14 March 2024
 # Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: need to have run 01-download_data.R

### 2024-03-14 scripts/03-test_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Tests the three data sets
 # Author: Mehrnoush Mohammadi & Quang Mai
-# Date: 2 March 2024
+# Date: 14 March 2024
 # Contact: mm.mohammadi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:

### 2024-03-14 outputs/paper/paper.qmd
Graphs moved to the data section

@@
 ---
-title: "Understanding the Landscape of Childlessness in the U.S.: A Data-Driven Exploration (1972-2022)"
+title: "Understanding the Landscape of Childlessness in the U.S."
+Subtitle: "An analysis of General Social Survey data (1972-2022)"
 author: 
@@
 date-format: long
-abstract: "This paper investigates U.S.
-childlessness trends (1972-2022) using data collected from the General Social Survey (GSS).
-The research covers several factors such as age, gender, education, social class, health conditions, marital status, and number of children.
-The analysis reveals rising childlessness across demographics, potentially linked to economic factors, education, and social norms.
-We explore the social and economic implications of this trend and emphasize the need for further research on childlessness in modern America."
+abstract: "We investigate U.S.
+childlessness trends between 1972 and 2022 using data from the General Social Survey.
+The research focuses on the relationship between childlessness and factors such as age, gender, education, social class, health conditions, marital status, and number of children.
+We find rising childlessness across demographics, potentially linked to economic factors, education, and social norms.
+We explore the social and economic implications of this trend and emphasize the need for further research on childlessness."
 format: pdf
@@
 table_nochild_age <- read_csv(file = "../data/clean_data_age.csv",show_col_types = FALSE)
-
 table_nochild_class <- read_csv(file = "../data/clean_data_class.csv",show_col_types = FALSE)
-
 table_nochild_degree <- read_csv(file = "../data/clean_data_degree.csv",show_col_types = FALSE)
-
 table_nochild_sex <- read_csv(file = "../data/clean_data_sex.csv",show_col_types = FALSE)
-
 table_nochild_marital <- read_csv(file = "../data/clean_data_marital.csv",show_col_types = FALSE)
-
 table_nochild_health <- read_csv(file = "../data/clean_data_health.csv",show_col_types = FALSE)
@@
 Traditionally viewed as a societal norm, having children now becomes a choice [@voluntaryChild].
-This paper looks into this phenomenon by examining various datasets on factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they affect childbearing rates over the past 50 years.
+We examine factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they affect childbearing rates over the past 50 years.
 
-Analyzing these factors offers a window into understanding not only the prevalence of childlessness within different segments of the population but also the potential drivers behind this trend.
+This offers a window into understanding not only the prevalence of childlessness within different segments of the population but also the potential drivers behind this trend.
 Existing research suggests a complex interplay of factors, including economic considerations, rising educational attainment, changing social norms related to career and family life, and individual preferences [@voluntaryChild].
 
-This paper aims to unpack these factors and explore their impact on the rise of childlessness in America.
+We explore the impact of these factors on the rise of childlessness in America.
 By analyzing the data, we seek to answer crucial questions: What are some important factors contributing to this trend?
@@
 Next, [Discussion] provides further insights and future areas of study.
-Finally, [Conclusion] concludes and summarizes our main findings.
+Finally, [Conclusion] summarizes our main findings.
 To complete the paper, [Appendix] clarifies how each variable within each dataset is measured.
@@
 This paper uses a multifaceted approach, drawing on various datasets to understand the complexities of childlessness in the United States.
-One primary source is the General Social Survey (GSS), a long-running and highly respected survey conducted by the National Opinion Research Center (NORC) at the University of Chicago [@gss], from 1972 to 2022.
+Our primary source is the General Social Survey (GSS), a long-running and highly respected survey conducted by the National Opinion Research Center (NORC) at the University of Chicago [@gss], from 1972 to 2022.
 
@@
 
-* Children: This variable captures the number of children ever born alive to the respondent, providing a direct measure of childlessness (having no children).
-* Year: This variable allows us to analyze trends in childlessness over time, encompassing the chosen time frame for the survey.
-* Age: This variable, categorized into relevant age groups (e.g., 18-34, 35-49, etc.), enables us to understand how childlessness manifests across different age demographics.
+* **Children:** This variable captures the number of children ever born alive to the respondent, providing a direct measure of childlessness (having no children).
+* **Year:** This variable allows us to analyze trends in childlessness over time, encompassing the chosen time frame for the survey.
+* **Age:** This variable, categorized into relevant age groups (e.g., 18-34, 35-49, etc.), enables us to understand how childlessness manifests across different age demographics.
 
@@
 
-### Exploring Socioeconomic Class and Educational Attainment
-To gain a deeper understanding of the factors influencing these trends, we look into additional data sets from NORC (@gss) examining the relationship between childlessness and socioeconomic class @tbl-nochild-class and educational attainment @tbl-nochild-degree.
+@fig-nochild-age shows a consistent rise in childlessness across all age groups in the United States from 1972 to 2022.
+It is important to acknowledge, however, that the data series has gaps for several years within this timeframe (1979, 1981, 1992, 1995, 1997, 1999, 2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017, and 2019).
+These missing data points can potentially limit the generalization of the observed trends and hinder a comprehensive understanding of childlessness over time.
+Despite these limitations, the available data reveals a steady increase in the percentage of individuals without children across all age groups.
+Notably, the most significant rise is observed among younger adults (18-34 and 35-49).
+As an example, the percentage of childless individuals in the 18-34 age group increased by approximately 42.70% from 1972 to 2022.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-age
+#| fig-cap: "Percentage of Childless Adults by Age Group in the US (1972-2020)"
+
+data_long_age <- tidyr::pivot_longer(table_nochild_age, -Age, names_to = "year", values_to = "percentage")
+
+ggplot(data_long_age, aes(x = as.integer(year), y = percentage, color = Age)) +
+  geom_point() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "solid", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Age Group", color = "Age Group") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+# Calculation of Percentage Increase in 18-34
+initial_value <- 46.07
+final_value <- 65.76
+percentage_increase <- ((final_value - initial_value) / initial_value) * 100
+
+#percentage_increase
+```
+
+## Exploring Socioeconomic Class and Educational Attainment
+To gain a deeper understanding of the factors influencing these trends, we look into additional data sets from NORC [@gss] examining the relationship between childlessness and socioeconomic class @tbl-nochild-class and educational attainment @tbl-nochild-degree.
 
@@
 
-* Socioeconomic Class: Data on socioeconomic class, often measured by income or occupation, will be utilized to understand the impact of economic factors on childbearing choices.
-* Educational Attainment: Data on educational attainment, such as the highest degree obtained, will be incorporated to explore the relationship between educational level and childbearing.
+* **Socioeconomic Class:** Data on socioeconomic class, often measured by income or occupation, will be utilized to understand the impact of economic factors on childbearing choices.
+* **Educational Attainment:** Data on educational attainment, such as the highest degree obtained, will be incorporated to explore the relationship between educational level and childbearing.
 
@@
 kable(table_nochild_class_subset, 
-      caption = "Childless Percentage of Population by Class Indentification and Selected Years", 
+      caption = "Childless Percentage of Population by Class Identification and Selected Years", 
       align = "c",
@@
 
+@fig-nochild-class depicts the gradual increase in the percentage of childless individuals across four social classes (lower class, working class, middle class, and upper class) in the United States from 1972 to 2022.
+The graph reveals an upward trend in childlessness for all classes over the five decades.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-class
+#| fig-cap: "Percentage Childless over Time by Class"
+
+# Pivot the data from wide to long format
+data_long_class <- pivot_longer(table_nochild_class, -`Class Identification`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_class$Year <- as.integer(data_long_class$Year)
+
+# Plot the data
+ggplot(data_long_class, aes(x = Year, y = Percentage, color = `Class Identification`)) +
+  geom_point() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "solid", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Class", color = "Breakdown Category") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+```
+
+People and households are divided into groups based on their socioeconomic status.
+The lower class experiences economic hardship, while the working class earns limited wages and faces job insecurity.
+The middle class has white-collar or professional occupations and enjoys a comfortable lifestyle.
+The upper class has substantial wealth and enjoys a privileged lifestyle with exclusive amenities [@pewresearch2022].
+
+While all classes show an increase in @fig-nochild-class, the data suggests the most significant rise occurred among the lower class, with their percentage of childless individuals nearly doubling from approximately 17.16% in 1972 to around 34.46% in 2022, marking a substantial increase of approximately 101.04% over the specified period.
+The working class also experienced a significant increase, from around 21.83% in 1972 to approximately 31.44% in 2022, indicating a rise of roughly 43.98%.
+Additionally, the middle class saw a slight increase, from approximately 28.24% in 1972 to about 30.09% in 2022, representing a rise of approximately 6.55%.
+Similarly, the upper class saw an increase, with their percentage of childless individuals climbing from approximately 20.24% in 1972 to about 27.87% in 2022, indicating an increase of roughly 37.71%.
+
 @tbl-nochild-degree explores the association between educational attainment and childlessness.
@@
 
+\newpage
 ```{r}
@@
 
-### Exploring Gender Differences
+@fig-nochild-degree presents a comparative analysis of childlessness across individuals with different levels of educational attainment (less than high school, high school and college degree or higher) in the United States from 1972 to 2022.
+The graph utilizes line trends to depict the percentage of childless individuals within each degree category over the five decades.
+
+The data reveals a distinct pattern across educational groups.
+Individuals with a high school diploma consistently exhibit the highest percentage of childlessness throughout the entire period, ranging from around 28.15% in 1972 to 34.56% in 2022.
+Conversely, the percentage of childless individuals with a college and higher degree has decreased from 32.85% in 1972 to 29.84% in 2022.
+The trend for individuals with a with less than a high school education increased from 16.97% in 1972 to 22.90% in 2022.
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-degree
+#| fig-cap: "Percentage Childless over Time by Level of Education"
+
+# Pivot the data from wide to long format
+data_long_degree <- pivot_longer(table_nochild_degree, -`Degree`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_degree$Year <- as.integer(data_long_degree$Year)
+
+# Plot the data
+ggplot(data_long_degree, aes(x = Year, y = Percentage, color = `Degree`)) +
+  geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Degree", color = "Degree") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+```
+
+## Exploring Gender Differences
 It is crucial to acknowledge potential gender disparities in childbearing experiences.
@@
 
-### Exploring Health Conditions and Marital Status
+@fig-nochild-gender explores gender differences in childlessness within the United States over a 50-year period (1972-2022).
+The graph utilizes line trends to depict the percentage of childless individuals categorized by gender (Male and Female).
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-gender
+#| fig-cap: "Percentage Childless over Time by Gender"
+
+# Pivot the data from wide to long format
+data_long_sex <- pivot_longer(table_nochild_sex, -`Gender`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_sex$Year <- as.integer(data_long_sex$Year)
+
+# Plot the data
+ggplot(data_long_sex, aes(x = Year, y = Percentage, color = `Gender`)) +
+  geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Gender", color = "Gender") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+```
+The data reveals a gradual increase in childlessness for both genders over the past five decades.
+However, females consistently exhibit a lower percentage of childlessness compared to males throughout the entire period.
+While the gap appears to be slightly narrowing over time, females still have a lower percentage by around 8 percentage points in 2022 (approximately 28.32% for females and 36.09% for males).
+
+## Exploring Health Conditions and Marital Status
 
 In addition, additional data sets from NORC [@gss] were also leveraged to understand the context of childlessness in America from 1972-2022.
-@tbl-nochild-health and @tbl-nochild-marital present an interesting connection between childlessness and health conditions and martial status.
+@tbl-nochild-health and @tbl-nochild-marital present an interesting connection between childlessness and health conditions and marital status.
 
@@
 
+@fig-nochild-health explores health differences in childlessness within the United States over a 50-year period (1972-2022).
+The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (excellent, good, fair, and poor).
+
+Overall, the data shows an increase in childlessness across all health levels and conditions approaching 2022.
+As mentioned, people with excellent health have the highest percentage of childlessness, having a 10.49% increase throughout the 50-year period (28.26% in 1972 to 38.75% in 2022).
+People with good and fair health exhibit a similar rate of childlessness in 2022, at about 30%.
+In particular, people with fair health marked; however, people with poor health, once having the lowest rate of childlessness (at 12.27% in 1972), now increases to 20.40% in 2022, which translates to a substantial 8.13% increase.
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-health
+#| fig-cap: "Percentage Childless over Time by Health Conditions"
+
+# Pivot the data from wide to long format
+data_long_health <- pivot_longer(table_nochild_health, -`Condition of health`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_health$Year <- as.integer(data_long_health$Year)
+
+# Plot the data
+ggplot(data_long_health, aes(x = Year, y = Percentage, color = `Condition of health`)) +
+  geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Health", color = "Condition of health") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+```
+
 @tbl-nochild-marital explores people's marital status and childlessness.
@@
 ```
+
+@fig-nochild-marital explores the marital status of people who are childless over a span of 50 years in the United States (1972-2022).
+The graph contains trend lines of varying marital statuses-Divorced, Married, Never married, Separated and Widowed-and how each has changed over the years in terms of childbearing status.
+
+The data demonstrates that a significant portion of people who have never married are childless.
+However, more and more people are reluctant to get married but still consider having children.
+This is observable in the childlessness percentage decrease of people who never marry, decreasing from 96.73% in 1972 to 72.53% in 2022.
+For people who are divorced, separated and widowed, this percentage fluctuates, displaying the complex dynamics between these relationship statuses and childbearing.
+For married people though, this childlessness percentage also decreases, from 10.51% in 1972 to 13.97% in 2022.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-marital
+#| fig-cap: "Percentage Childless over Time by Marital Status"
+
+# Pivot the data from wide to long format
+data_long_marital <- pivot_longer(table_nochild_marital, -`Marital status`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_marital$Year <- as.integer(data_long_marital$Year)
+
+# Plot the data
+ggplot(data_long_marital, aes(x = Year, y = Percentage, color = `Marital status`)) +
+  geom_point() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "solid", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Marital Status", color = "Marital status") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+```
+
+The figures presented in this paper were generated using data from the [@gss].
+These figures visually represent the key findings related to the trends and patterns in childlessness across different ages, social classes, educational attainment levels, and genders in the United States over the past five decades (1972-2022).
+It is important to remember that these figures solely present descriptive data and do not establish causal relationships.
+The Discussion section will further analyze the observed trends and explore the broader implications of these findings.
 
@@
 
-@fig-nochild-age shows a consistent rise in childlessness across all age groups in the United States from 1972 to 2022.
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-age
-#| fig-cap: "Percentage of Childless Adults by Age Group in the US (1972-2020)"
-
-data_long_age <- tidyr::pivot_longer(table_nochild_age, -Age, names_to = "year", values_to = "percentage")
-
-ggplot(data_long_age, aes(x = as.integer(year), y = percentage, color = Age)) +
-  geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Age Group", color = "Age Group") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-# Calculation of Percentage Increase in 18-34
-initial_value <- 46.07
-final_value <- 65.76
-percentage_increase <- ((final_value - initial_value) / initial_value) * 100
-
-#percentage_increase
-```
-
-\newpage
-
-It is important to acknowledge, however, that the data series has gaps for several years within this timeframe (1979, 1981, 1992, 1995, 1997, 1999, 2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017, and 2019).
-These missing data points can potentially limit the generalization of the observed trends and hinder a comprehensive understanding of childlessness over time.
-Despite these limitations, the available data reveals a steady increase in the percentage of individuals without children across all age groups.
-Notably, the most significant rise is observed among younger adults (18-34 and 35-49).
-As an example, the percentage of childless individuals in the 18-34 age group increased by approximately 42.70% from 1972 to 2022.
-
-@fig-nochild-class depicts the gradual increase in the percentage of childless individuals across four social classes (lower class, working class, middle class, and upper class) in the United States from 1972 to 2022.
-The graph reveals an upward trend in childlessness for all classes over the five decades.
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-class
-#| fig-cap: "Percentage Childless over Time by Class"
-
-# Pivot the data from wide to long format
-data_long_class <- pivot_longer(table_nochild_class, -`Class Identification`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_class$Year <- as.integer(data_long_class$Year)
-
-# Plot the data
-ggplot(data_long_class, aes(x = Year, y = Percentage, color = `Class Identification`)) +
-  geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Class", color = "Breakdown Category") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-```
-
-People and households are divided into groups based on their socioeconomic status.
-The lower class experiences economic hardship, while the working class earns limited wages and faces job insecurity.
-The middle class has white-collar or professional occupations and enjoys a comfortable lifestyle.
-The upper class has substantial wealth and enjoys a privileged lifestyle with exclusive amenities [@pewresearch2022].
-
-While all classes show an increase in @fig-nochild-class, the data suggests the most significant rise occurred among the lower class, with their percentage of childless individuals nearly doubling from approximately 17.16% in 1972 to around 34.46% in 2022, marking a substantial increase of approximately 101.04% over the specified period.
-The working class also experienced a significant increase, from around 21.83% in 1972 to approximately 31.44% in 2022, indicating a rise of roughly 43.98%.
-Additionally, the middle class saw a slight increase, from approximately 28.24% in 1972 to about 30.09% in 2022, representing a rise of approximately 6.55%.
-Similarly, the upper class saw an increase, with their percentage of childless individuals climbing from approximately 20.24% in 1972 to about 27.87% in 2022, indicating an increase of roughly 37.71%.
-
-@fig-nochild-degree presents a comparative analysis of childlessness across individuals with different levels of educational attainment (less than high school, high school and college degree or higher) in the United States from 1972 to 2022.
-The graph utilizes line trends to depict the percentage of childless individuals within each degree category over the five decades.
-
-The data reveals a distinct pattern across educational groups.
-Individuals with a high school diploma consistently exhibit the highest percentage of childlessness throughout the entire period, ranging from around 28.15% in 1972 to 34.56% in 2022.
-Conversely, the percentage of childless individuals with a college and higher degree has decreased from 32.85% in 1972 to 29.84% in 2022.
-The trend for individuals with a with less than a high school education increased from 16.97% in 1972 to 22.90% in 2022.
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-degree
-#| fig-cap: "Percentage Childless over Time by Level of Education"
-
-# Pivot the data from wide to long format
-data_long_degree <- pivot_longer(table_nochild_degree, -`Degree`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_degree$Year <- as.integer(data_long_degree$Year)
-
-# Plot the data
-ggplot(data_long_degree, aes(x = Year, y = Percentage, color = `Degree`)) +
-  geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Degree", color = "Degree") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-```
-
-\newpage
-
-@fig-nochild-gender explores gender differences in childlessness within the United States over a 50-year period (1972-2022).
-The graph utilizes line trends to depict the percentage of childless individuals categorized by gender (Male and Female).
-
-The data reveals a gradual increase in childlessness for both genders over the past five decades.
-However, females consistently exhibit a lower percentage of childlessness compared to males throughout the entire period.
-While the gap appears to be slightly narrowing over time, females still have a lower percentage by around 8 percentage points in 2022 (approximately 28.32% for females and 36.09% for males).
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-gender
-#| fig-cap: "Percentage Childless over Time by Gender"
-
-# Pivot the data from wide to long format
-data_long_sex <- pivot_longer(table_nochild_sex, -`Gender`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_sex$Year <- as.integer(data_long_sex$Year)
-
-# Plot the data
-ggplot(data_long_sex, aes(x = Year, y = Percentage, color = `Gender`)) +
-  geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Gender", color = "Gender") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-```
-
-@fig-nochild-health explores health differences in childlessness within the United States over a 50-year period (1972-2022).
-The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (excellent, good, fair, and poor).
-
-Overall, the data shows an increase in childlessness across all health levels and conditions approaching 2022.
-As mentioned, people with excellent health have the highest percentage of childlessness, having a 10.49% increase throughout the 50-year period (28.26% in 1972 to 38.75% in 2022).
-People with good and fair health exhibit a similar rate of childlessness in 2022, at about 30%.
-In particular, people with fair health marked; however, people with poor health, once having the lowest rate of childlessness (at 12.27% in 1972), now increases to 20.40% in 2022, which translates to a substantial 8.13% increase.
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-health
-#| fig-cap: "Percentage Childless over Time by Health Conditions"
-
-# Pivot the data from wide to long format
-data_long_health <- pivot_longer(table_nochild_health, -`Condition of health`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_health$Year <- as.integer(data_long_health$Year)
-
-# Plot the data
-ggplot(data_long_health, aes(x = Year, y = Percentage, color = `Condition of health`)) +
-  geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Health", color = "Condition of health") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-```
-
-\newpage
-
-@fig-nochild-marital explores the marital status of people who are childless over a span of 50 years in the United States (1972-2022).
-The graph contains trend lines of varying marital statuses-Divorced, Married, Never married, Separated and Widowed-and how each has changed over the years in terms of childbearing status.
-
-The data demonstrates that a significant portion of people who have never married are childless.
-However, more and more people are reluctant to get married but still consider having children.
-This is observable in the childlessness percentage decrease of people who never marry, decreasing from 96.73% in 1972 to 72.53% in 2022.
-For people who are divorced, separated and widowed, this percentage fluctuates, displaying the complex dynamics between these relationship statuses and childbearing.
-For married people though, this childlessness percentage also decreases, from 10.51% in 1972 to 13.97% in 2022.
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-marital
-#| fig-cap: "Percentage Childless over Time by Marital Status"
-
-# Pivot the data from wide to long format
-data_long_marital <- pivot_longer(table_nochild_marital, -`Marital status`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_marital$Year <- as.integer(data_long_marital$Year)
-
-# Plot the data
-ggplot(data_long_marital, aes(x = Year, y = Percentage, color = `Marital status`)) +
-  geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Marital Status", color = "Marital status") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-```
-
-The figures presented in this paper were generated using data from the [@gss].
-These figures visually represent the key findings related to the trends and patterns in childlessness across different ages, social classes, educational attainment levels, and genders in the United States over the past five decades (1972-2022).
-It is important to remember that these figures solely present descriptive data and do not establish causal relationships.
-The Discussion section will further analyze the observed trends and explore the broader implications of these findings.
-
 ```{r}
@@
 #| label: fig-nochild-avg-marital
-#| fig-cap: "Average Childless Percentage of Population over Time by Marital Status"
+#| fig-cap: "Average Percentage Childless over Time by Marital Status"
 
 # Group the data by Year and calculate the average of Percentage for each year
-avg_data <- data_long_marital |>
+avg_data_marital <- data_long_marital |>
   group_by(Year) |>
@@
 # Print the average data
-# print(avg_data)
+# print(avg_data_marital)
 
 # Plot the data
-ggplot(avg_data, aes(x = Year, y = avg_Percentage)) +
+ggplot(avg_data_marital, aes(x = Year, y = avg_Percentage)) +
   geom_line() +
@@
   labs(x = "Year", y = "Average Percentage of Population", title = "Average Childless Percentage of Population over Time by Marital Status") +
-  theme_minimal()
+  theme_minimal() +
+  theme(text = element_text(size = 9))
 
 ```
+
 ```{r}
@@
 # Group the data by Year and calculate the average of Percentage for each year
-avg_data <- data_long_class |>
+avg_data_class <- data_long_class |>
   group_by(Year) |>
@@
 # Print the average data
-# print(avg_data)
+# print(avg_data_class)
 
 # Plot the data
-ggplot(avg_data, aes(x = Year, y = avg_Percentage)) +
+ggplot(avg_data_class, aes(x = Year, y = avg_Percentage)) +
   geom_line() +
@@
   labs(x = "Year", y = "Average Percentage of Population", title = "Average Percentage Childless over Time by Class") +
-  theme_minimal()
+  theme_minimal() +
+  theme(text = element_text(size = 9))
 
@@
 # Group the data by Year and calculate the average of Percentage for each year
-avg_data <- data_long_sex |>
+avg_data_sex <- data_long_sex |>
   group_by(Year) |>
@@
 # Print the average data
-# print(avg_data)
+# print(avg_data_sex)
 
 # Plot the data
-ggplot(avg_data, aes(x = Year, y = avg_Percentage)) +
+ggplot(avg_data_sex, aes(x = Year, y = avg_Percentage)) +
   geom_line() +
@@
   labs(x = "Year", y = "Average Percentage of Population", title = "Average Percentage Childless over Time by Gender") +
-  theme_minimal()
+  theme_minimal() +
+  theme(text = element_text(size = 9))
 
@@
 # Group the data by Year and calculate the average of Percentage for each year
-avg_data <- data_long_degree |>
+avg_data_degree <- data_long_degree |>
   group_by(Year) |>
@@
 # Print the average data
-# print(avg_data)
+# print(avg_data_degree)
 
 # Plot the data
-ggplot(avg_data, aes(x = Year, y = avg_Percentage)) +
+ggplot(avg_data_degree, aes(x = Year, y = avg_Percentage)) +
   geom_line() +
@@
   labs(x = "Year", y = "Average Percentage of Population", title = "Average Percentage Childless over Time by Degree") +
-  theme_minimal()
+  theme_minimal() +
+  theme(text = element_text(size = 9))
 
 ```
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: fig-nochild-avg-degree-class
+#| fig-cap: "Average Percentage Childless over Time by Degree and Class"
+
+# Merge the two datasets based on the Year column
+merged_data <- merge(avg_data_degree, avg_data_class, by = "Year", all = TRUE, suffixes = c("_degree", "_class"))
+
+# Calculate the sum of avg_Percentage for each year
+merged_data <- merged_data |>
+  mutate(sum_Percentage = (coalesce(avg_Percentage_degree, 0) + coalesce(avg_Percentage_class, 0))/2)
+
+# Print the merged and summed data
+# print(merged_data)
+
+# Plot the merged and summed data
+ggplot(merged_data, aes(x = Year, y = sum_Percentage)) +
+  geom_line() +
+  geom_point() +
+  labs(x = "Year", y = "Sum of Average Percentage of Population", title = "Sum of Average Percentage Childless over Time by Degree and Class") +
+  theme_minimal() +
+  theme(text = element_text(size = 9))
+```
+
+\newpage
 
@@
 The findings presented in this paper reveal a consistent and concerning trend of rising childlessness across various demographics in the United States.
-As Figure 1 illustrates, the percentage of individuals without children has steadily increased across all age groups over the past 50 years, with the most significant rise observed among younger individuals (18-34 and 35-49).
+As @fig-nochild-age illustrates, the percentage of individuals without children has steadily increased across all age groups over the past 50 years, with the most significant rise observed among younger individuals (18-34 and 35-49).
 

### 2024-03-18 outputs/paper/paper.qmd
finalize paper

@@
 title: "Understanding the Landscape of Childlessness in the U.S."
-Subtitle: "An analysis of General Social Survey data (1972-2022)"
+subtitle: "An analysis of General Social Survey data (1972-2022)"
 author: 
@@
 table_nochild_health <- read_csv(file = "../data/clean_data_health.csv",show_col_types = FALSE)
+table_nochild_overall <- read_csv(file =
+"../data/nochild_overall.csv",show_col_types = FALSE)
 
+table_nochild_overall <- table_nochild_overall |>
+  drop_na() |>
+  mutate_at(vars(-1), as.numeric)
 ```
@@
 
-This offers a window into understanding not only the prevalence of childlessness within different segments of the population but also the potential drivers behind this trend.
+These helps us understand not only the prevalence of childlessness within different segments of the population but also the potential drivers behind this trend.
 Existing research suggests a complex interplay of factors, including economic considerations, rising educational attainment, changing social norms related to career and family life, and individual preferences [@voluntaryChild].
@@
 
-Structurally, in [Introduction], our paper starts with a brief dissection of the nature of childlessness in America and the contributing factors.
+Our estimand is the relationship between age, gender, socio-economic class, educational attainment, health status, marital status and the rate of childlessness in the United States from 1972 to 2022.
+Through our analysis, we found that childless rate is consistently on the rise from 1972 to 2022, with correlation coefficient of 0.73 between childless percentage and time (or a 6.99% increase from 1972 to 2022).
+Factors that directly contribute to this rise are higher educational attainment, higher socio-economic class,lower gender disparity, better health status and dissolution of marriages through divorces, separation and being widowed.
+
+To further understand why these causes explain rising childlessness, our paper starts with a brief dissection of the nature of childlessness in America and the contributing factors in [Introduction].
 Subsequently, in [Data] and [Result], we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.
@@
 Finally, [Conclusion] summarizes our main findings.
-To complete the paper, [Appendix] clarifies how each variable within each dataset is measured.
+To complete the paper, [Appendix] clarifies how each variable within each dataset is measured and the survey that we've generated based off the dataset.
 
@@
 
+## Measurement
+
+Each variable used—age, class, educational attainment, health, marital status, and gender—is  quantified through direct questioning to accurately capture respondents' backgrounds.
+For instance, the General Social Survey (GSS) obtains information on marital status through questions such as, "Are you currently married, widowed, divorced, separated, or have you never been married?", with responses categorized into predefined groups for further analysis [@gssCodebook].
+On ther other hand, instead of relying on income sources, socio-economic class is subjectively assessed based on the respondents' answers to this question, "If you were asked to use one of four names for your social class, which would you say you belong in: the lower class, the working class, the middle class, or the upper class?".
+Some are continuous variable, like age.
+With respondents' various age groups documented, we can understand the generational disparities in childbearing patterns.
+Systematic and data-driven, these measurement methods ensures that all data faithfully reflects the American population's childbearing tendencies.
+
 ## Source Data
 
-The 2022 iteration of the GSS, for the first time in its history, incorporated a combination of in-person, web self-administered, and phone interview methods.
+The 2022 iteration of the GSS, for the first time in its history, incorporated a combination of in-person, web self-administered, and phone interview methods in response to COVID-19.
 This expansion in data collection methods offers a potentially richer and more diverse sample.
-The GSS team is transparent about these limitations and welcomes feedback from the research community to continuously improve their methodology.
+As a result, after implementing these changes, the GSS randomly assigned respondents to various survey modes to examine response patterns and data accuracy.
+Unsurprisingly, compared to last year's 17%, 94.2% of participants opted for a new survey mode, which is a significant increase of 50.5% in the overall response rate.
+This emphasizes the effectiveness of employing multimodal survey strategies to boost survey participation.
+
+However, this multimodal survey approach also presents several challenges in collecting respondents' opinions on topics such as childbirth and childbearing tendencies, given the private nature of these subjects.
+With the easy formatting and non-required nature of these surveys, biases like acquiescence bias (only positive responses) might occur because respondents aren't confronted in a face-to-face settings, where many will just opt a neutral response [@webSurvey].
+Therefore, we carefully select various variables to to ensure the authenticity of the public's background on the topic of childlessness, avoiding having the data being too influenced by the changes in surveying methods.
+
+## Exploring Age and Generational Differences
 
@@
 
-\newpage
+Thus, there is a consistent rise in childlessness across all age groups in the United States from 1972 to 2022.
+Notably, the most significant rise is observed among younger adults (18-34 and 35-49).
+It is important to acknowledge, however, that the data series has gaps for several years within this timeframe (1979, 1981, 1992, 1995, 1997, 1999, 2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017, and 2019).
+These missing data points can potentially limit the generalization of the observed trends and hinder a comprehensive understanding of childlessness over time.
 
@@
 
-@fig-nochild-age shows a consistent rise in childlessness across all age groups in the United States from 1972 to 2022.
-It is important to acknowledge, however, that the data series has gaps for several years within this timeframe (1979, 1981, 1992, 1995, 1997, 1999, 2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017, and 2019).
-These missing data points can potentially limit the generalization of the observed trends and hinder a comprehensive understanding of childlessness over time.
-Despite these limitations, the available data reveals a steady increase in the percentage of individuals without children across all age groups.
-Notably, the most significant rise is observed among younger adults (18-34 and 35-49).
-As an example, the percentage of childless individuals in the 18-34 age group increased by approximately 42.70% from 1972 to 2022.
-
 ```{r}
 #| echo: false
-#| include: true
+#| include: false
 #| warning: false
 #| label: fig-nochild-age
-#| fig-cap: "Percentage of Childless Adults by Age Group in the US (1972-2020)"
+#| fig-cap: "Childless Percentage in the US by Age Group (1972-2020)"
 
@@
 ggplot(data_long_age, aes(x = as.integer(year), y = percentage, color = Age)) +
-  geom_point() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "solid", size = 1.2) +  # Add running average trend line
+   geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
   labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Age Group", color = "Age Group") +
@@
 
+## Exploring Gender Differences
+It is crucial to acknowledge potential gender disparities in childbearing experiences.
+A separate analysis examining childlessness by gender (presented in @tbl-nochild-sex) aims to provide a comprehensive understanding by considering the unique experiences of men and women regarding childlessness.
+Data on gender is included from NORC [@gss] to allow for this investigation.
+
+The data reveals a gradual increase in childlessness for both genders over the past five decades.
+However, females consistently exhibit a lower percentage of childlessness compared to males throughout the entire period.
+While the gap appears to be slightly narrowing over time, females still have a lower percentage by around 8 percentage points in 2022 (approximately 28.32% for females and 36.09% for males).
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-sex
+
+# Selected years to show better on the paper
+random_years <- c("Gender", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_sex_subset <- table_nochild_sex |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_sex_subset, 
+      caption = "Childless Percentage of Population by Gender and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
+```{r}
+#| echo: false
+#| include: false
+#| warning: false
+#| label: fig-nochild-gender
+#| fig-cap: "Childless Percentage in the US over Time by Gender (1972-2020)"
+
+# Pivot the data from wide to long format
+data_long_sex <- pivot_longer(table_nochild_sex, -`Gender`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_sex$Year <- as.integer(data_long_sex$Year)
+
+# Plot the data
+ggplot(data_long_sex, aes(x = Year, y = Percentage, color = `Gender`)) +
+  geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Gender", color = "Gender") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+```
+
 ## Exploring Socioeconomic Class and Educational Attainment
-To gain a deeper understanding of the factors influencing these trends, we look into additional data sets from NORC [@gss] examining the relationship between childlessness and socioeconomic class @tbl-nochild-class and educational attainment @tbl-nochild-degree.
+To gain a deeper understanding of the factors influencing these trends, we look into additional data sets from NORC [@gss] examining the relationship between childlessness and socioeconomic class (@tbl-nochild-class) and educational attainment (@tbl-nochild-degree).
+
+```{r}
+#| include: true
+#| label: fig-nochild-class-edu
+#| fig-cap: Childless Percentage in the US by Class and Educational Attainment (1972-2020)
+#| fig-subcap: [Childless percentage by class, Childless percentage by education]
+#| layout-ncol: 2
+#| warning: false
+#| message: false
+#| echo: false
+#| fig-pos: "t"
+
+# Pivot the data from wide to long format
+data_long_class <- pivot_longer(table_nochild_class, -`Class Identification`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_class$Year <- as.integer(data_long_class$Year)
+
+# Plot the data
+ggplot(data_long_class, aes(x = Year, y = Percentage, color = `Class Identification`)) +
+   geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Class", color = "Breakdown Category") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+# Pivot the data from wide to long format
+data_long_degree <- pivot_longer(table_nochild_degree, -`Degree`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_degree$Year <- as.integer(data_long_degree$Year)
+
+# Plot the data
+ggplot(data_long_degree, aes(x = Year, y = Percentage, color = `Degree`)) +
+  geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Degree", color = "Degree") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+```
 
@@
 
-@tbl-nochild-class reveals a potential link between socioeconomic class and childlessness.
-The data suggests that individuals in the lower and working classes exhibit higher percentages of childlessness compared to those in the middle and upper classes.
-This pattern might be attributed to various factors, such as financial constraints, job insecurity, or limited access to childcare resources experienced by individuals in lower socioeconomic classes.
+### Overview of Socioeconomic Class's Impact on Childlessness
 
@@
 
-@fig-nochild-class depicts the gradual increase in the percentage of childless individuals across four social classes (lower class, working class, middle class, and upper class) in the United States from 1972 to 2022.
+@tbl-nochild-class reveals a potential link between socioeconomic class and childlessness.
+To visualize the data, @fig-nochild-class-edu depicts the gradual increase in the percentage of childless individuals across four social classes (lower class, working class, middle class, and upper class) in the United States from 1972 to 2022.
 The graph reveals an upward trend in childlessness for all classes over the five decades.
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-class
-#| fig-cap: "Percentage Childless over Time by Class"
-
-# Pivot the data from wide to long format
-data_long_class <- pivot_longer(table_nochild_class, -`Class Identification`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_class$Year <- as.integer(data_long_class$Year)
-
-# Plot the data
-ggplot(data_long_class, aes(x = Year, y = Percentage, color = `Class Identification`)) +
-  geom_point() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "solid", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Class", color = "Breakdown Category") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-```
 
@@
 
-While all classes show an increase in @fig-nochild-class, the data suggests the most significant rise occurred among the lower class, with their percentage of childless individuals nearly doubling from approximately 17.16% in 1972 to around 34.46% in 2022, marking a substantial increase of approximately 101.04% over the specified period.
-The working class also experienced a significant increase, from around 21.83% in 1972 to approximately 31.44% in 2022, indicating a rise of roughly 43.98%.
-Additionally, the middle class saw a slight increase, from approximately 28.24% in 1972 to about 30.09% in 2022, representing a rise of approximately 6.55%.
-Similarly, the upper class saw an increase, with their percentage of childless individuals climbing from approximately 20.24% in 1972 to about 27.87% in 2022, indicating an increase of roughly 37.71%.
+\newpage
 
-@tbl-nochild-degree explores the association between educational attainment and childlessness.
-Interestingly, the data indicates a nuanced relationship.
-While individuals with higher educational degrees (college+) display a slightly higher percentage of childlessness compared to those with lower degrees, the difference is not as substantial as observed with socioeconomic class.
-This suggests that educational attainment might play a complex role in childbearing decisions, potentially influenced by factors like career aspirations, financial considerations, and personal preferences.
+### Overview of Educational Attainment's Impact on Childlessness
 
-\newpage
+Further, @tbl-nochild-degree explores the association between educational attainment and childlessness.
+@fig-nochild-class-edu illustrates this data, presenting a comparative analysis of childlessness across individuals with different levels of educational attainment (less than high school, high school and college degree or higher) in the United States from 1972 to 2022.
+The graph utilizes line trends to depict the percentage of childless individuals within each degree category over the five decades.
+
 ```{r}
@@
 
-@fig-nochild-degree presents a comparative analysis of childlessness across individuals with different levels of educational attainment (less than high school, high school and college degree or higher) in the United States from 1972 to 2022.
-The graph utilizes line trends to depict the percentage of childless individuals within each degree category over the five decades.
-
-The data reveals a distinct pattern across educational groups.
-Individuals with a high school diploma consistently exhibit the highest percentage of childlessness throughout the entire period, ranging from around 28.15% in 1972 to 34.56% in 2022.
-Conversely, the percentage of childless individuals with a college and higher degree has decreased from 32.85% in 1972 to 29.84% in 2022.
-The trend for individuals with a with less than a high school education increased from 16.97% in 1972 to 22.90% in 2022.
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-degree
-#| fig-cap: "Percentage Childless over Time by Level of Education"
-
-# Pivot the data from wide to long format
-data_long_degree <- pivot_longer(table_nochild_degree, -`Degree`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_degree$Year <- as.integer(data_long_degree$Year)
-
-# Plot the data
-ggplot(data_long_degree, aes(x = Year, y = Percentage, color = `Degree`)) +
-  geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Degree", color = "Degree") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-```
-
-## Exploring Gender Differences
-It is crucial to acknowledge potential gender disparities in childbearing experiences.
-A separate analysis examining childlessness by gender (presented in @tbl-nochild-sex) aims to provide a comprehensive understanding by considering the unique experiences of men and women regarding childlessness.
-Data on gender is included from NORC [@gss] to allow for this investigation.
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: tbl-nochild-sex
-
-# Selected years to show better on the paper
-random_years <- c("Gender", "1972", "1980", "1990", "2000", "2010", "2022")
-
-# Subset the data with only the selected columns
-table_nochild_sex_subset <- table_nochild_sex |>
-  select(all_of(random_years))
-
-# Print the table with formatting
-kable(table_nochild_sex_subset, 
-      caption = "Childless Percentage of Population by Gender and Selected Years", 
-      align = "c",
-      col.names = random_years,
-      row.names = FALSE) |>
-  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
-
-```
-
-@fig-nochild-gender explores gender differences in childlessness within the United States over a 50-year period (1972-2022).
-The graph utilizes line trends to depict the percentage of childless individuals categorized by gender (Male and Female).
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-gender
-#| fig-cap: "Percentage Childless over Time by Gender"
-
-# Pivot the data from wide to long format
-data_long_sex <- pivot_longer(table_nochild_sex, -`Gender`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_sex$Year <- as.integer(data_long_sex$Year)
-
-# Plot the data
-ggplot(data_long_sex, aes(x = Year, y = Percentage, color = `Gender`)) +
-  geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Gender", color = "Gender") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-```
-The data reveals a gradual increase in childlessness for both genders over the past five decades.
-However, females consistently exhibit a lower percentage of childlessness compared to males throughout the entire period.
-While the gap appears to be slightly narrowing over time, females still have a lower percentage by around 8 percentage points in 2022 (approximately 28.32% for females and 36.09% for males).
-
 ## Exploring Health Conditions and Marital Status
@@
 
+```{r}
+#| include: true
+#| label: fig-nochild-health-marital
+#| fig-cap: Childless Percentage in the US over Time by Marital and Health Status (1972-2020)
+#| fig-subcap: [Childless percentage by health, Childless percentage by marital status]
+#| layout-ncol: 2
+#| warning: false
+#| message: false
+#| echo: false
+#| fig-pos: "t"
+
+# Pivot the data from wide to long format
+data_long_health <- pivot_longer(table_nochild_health, -`Condition of health`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_health$Year <- as.integer(data_long_health$Year)
+
+# Plot the data
+ggplot(data_long_health, aes(x = Year, y = Percentage, color = `Condition of health`)) +
+  geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Health", color = "Condition of health") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+# Pivot the data from wide to long format
+data_long_marital <- pivot_longer(table_nochild_marital, -`Marital status`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_marital$Year <- as.integer(data_long_marital$Year)
+
+# Plot the data
+ggplot(data_long_marital, aes(x = Year, y = Percentage, color = `Marital status`)) +
+  geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Marital Status", color = "Marital status") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+```
+
+\newpage
+### Overview of Health Status's Impact on Childlessness
+
 @tbl-nochild-health uncovers the relationship between health conditions and childlessness.
-Intriguingly, those with 'excellent' and 'good' health are substantially more likely to be childless than those with 'fair' and 'poor' health.
-Several factors are attributable to this fact, but foremost, career, lifestyle differences and financial considerations might come into play.
-However, more data is needed to clarify the context of this trend, especially coupling it with variables like educational attainment or socioeconomic class.
-In addition, sampling bias might factor into this trend, where it's assumable that people with excellent health are often the younger population and those with less desirable health are older.
+Accordingly, the first graph of @fig-nochild-health-marital explores the health differences in people who are childless within the United States over a 50-year period (1972-2022).
+The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (excellent, good, fair, and poor).
 
@@
 
-@fig-nochild-health explores health differences in childlessness within the United States over a 50-year period (1972-2022).
-The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (excellent, good, fair, and poor).
-
-Overall, the data shows an increase in childlessness across all health levels and conditions approaching 2022.
-As mentioned, people with excellent health have the highest percentage of childlessness, having a 10.49% increase throughout the 50-year period (28.26% in 1972 to 38.75% in 2022).
-People with good and fair health exhibit a similar rate of childlessness in 2022, at about 30%.
-In particular, people with fair health marked; however, people with poor health, once having the lowest rate of childlessness (at 12.27% in 1972), now increases to 20.40% in 2022, which translates to a substantial 8.13% increase.
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-health
-#| fig-cap: "Percentage Childless over Time by Health Conditions"
-
-# Pivot the data from wide to long format
-data_long_health <- pivot_longer(table_nochild_health, -`Condition of health`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_health$Year <- as.integer(data_long_health$Year)
-
-# Plot the data
-ggplot(data_long_health, aes(x = Year, y = Percentage, color = `Condition of health`)) +
-  geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Health", color = "Condition of health") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-```
+### Overview of Marital Status's Impact on Childlessness
 
 @tbl-nochild-marital explores people's marital status and childlessness.
-Trends of childlessness vary with this dataset.
-Unsurprisingly, people who are never married have the highest percentage of childlessness.
-However, those who are divorced, widowed and separated have fluctuating percentages of childlessness, which is understandable due to the complex emotional nature of these marital statuses.
+@fig-nochild-health-marital then illustrates this data, exploring the marital status of people who are childless over a span of 50 years in the United States (1972-2022).
+The graph contains trend lines of varying marital statuses-Divorced, Married, Never married, Separated and Widowed-and how each has changed over the years in terms of childbearing status.
+
+\newpage
 
@@
 
-@fig-nochild-marital explores the marital status of people who are childless over a span of 50 years in the United States (1972-2022).
-The graph contains trend lines of varying marital statuses-Divorced, Married, Never married, Separated and Widowed-and how each has changed over the years in terms of childbearing status.
-
-The data demonstrates that a significant portion of people who have never married are childless.
-However, more and more people are reluctant to get married but still consider having children.
-This is observable in the childlessness percentage decrease of people who never marry, decreasing from 96.73% in 1972 to 72.53% in 2022.
-For people who are divorced, separated and widowed, this percentage fluctuates, displaying the complex dynamics between these relationship statuses and childbearing.
-For married people though, this childlessness percentage also decreases, from 10.51% in 1972 to 13.97% in 2022.
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-marital
-#| fig-cap: "Percentage Childless over Time by Marital Status"
-
-# Pivot the data from wide to long format
-data_long_marital <- pivot_longer(table_nochild_marital, -`Marital status`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_marital$Year <- as.integer(data_long_marital$Year)
-
-# Plot the data
-ggplot(data_long_marital, aes(x = Year, y = Percentage, color = `Marital status`)) +
-  geom_point() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "solid", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Marital Status", color = "Marital status") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-```
-
-The figures presented in this paper were generated using data from the [@gss].
-These figures visually represent the key findings related to the trends and patterns in childlessness across different ages, social classes, educational attainment levels, and genders in the United States over the past five decades (1972-2022).
-It is important to remember that these figures solely present descriptive data and do not establish causal relationships.
-The Discussion section will further analyze the observed trends and explore the broader implications of these findings.
-
 ## Data Limitations
@@
 While the General Social Survey provides valuable data on childlessness trends, there are gaps in data availability for certain years.
+In summary, out of 68596 survey entrees, there are 250 missing cases.
 This creates a limitation in our ability to present a completely continuous picture of trends across the entire 50-year time frame.
@@
 
+## Health Conditions and Marital Status Impact on Childlessness
+
+Overall, the data shows an increase in childlessness across all health levels and conditions approaching 2020.
+As mentioned, @fig-nochild-health-marital shows that people with excellent health have the highest percentage of childlessness, having a 10.49% increase throughout the 50-year period (28.26% in 1972 to 38.75% in 2022).
+People with good and fair health exhibit a similar rate of childlessness in 2022, at about 30%.
+However, people with poor health, once having the lowest rate of childlessness (at 12.27% in 1972), now increases to 20.40% in 2022, which translates to a substantial 8.13% increase.
+Overall, on average, across all health levels from 1972-2022, @fig-avg-one shows that percentage of childlessness increases at about 6.836%.
+
+Unintuitively, those with 'excellent' and 'good' health are substantially more likely to be childless than those with 'fair' and 'poor' health.
+Several factors are attributable to this fact, but foremost, career, lifestyle differences and financial considerations might come into play.
+However, more data is needed to clarify the context of this trend, especially coupling it with variables like educational attainment or socioeconomic class.
+In addition, sampling bias might factor into this trend, where it's assumable that people with excellent health are often the younger population and those with less desirable health are older.
+
+As for marital status, @fig-nochild-health-marital demonstrates that a significant portion of people who have never married are childless.
+However, more and more people are reluctant to get married but still have children.
+This is observable in the childlessness percentage decrease of people who never marry, decreasing from 96.73% in 1972 to 72.53% in 2022.
+For people who are divorced, separated and widowed, this percentage is on the rise fluctuates, displaying the complex dynamics between these relationship statuses and childbearing.
+For married people though, this childlessness percentage also decreases, from 10.51% in 1972 to 13.97% in 2022.
+Overall, on average, across all marital status from 1972-2022, @fig-avg-one shows that percentage of childlessness decreases at about 7.39%.
+
+Trends of childlessness vary with this dataset.
+Unsurprisingly, people who are never married have the highest percentage of childlessness.
+However, those who are divorced, widowed and separated have fluctuating but similarly increasing percentages of childlessness, which is understandable due to the complex emotional nature of these marital statuses.
+The overall trend of people who are in a relationship or are married having more children (7.39% decrease of childlessness) are attributed to the higher quality of life, and other factors.
+
 ```{r}
+#| include: true
+#| label: fig-avg-one
+#| fig-cap: Average Childless Percentage per Year in the US (1972-2022)
+#| fig-subcap: [Childless percentage by health, Childless percentage by marital status]
+#| layout-ncol: 2
+#| warning: false
+#| message: false
 #| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-avg-marital
-#| fig-cap: "Average Percentage Childless over Time by Marital Status"
+#| fig-pos: "t"
+
+# Group the data by Year and calculate the average of Percentage for each year
+avg_data_health <- data_long_health |>
+  group_by(Year) |>
+  summarise(Sum_Percentage = sum(Percentage, na.rm = TRUE)) |>
+  mutate(avg_Percentage = Sum_Percentage / 5)
+
+# Print the average data
+# print(avg_data_health)
+
+# Plot the data
+ggplot(avg_data_health, aes(x = Year, y = avg_Percentage)) +
+  geom_line() +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE, color = "blue") +
+  labs(x = "Year", y = "Average Percentage per Year") +
+  theme_minimal() +
+  theme(text = element_text(size = 9))
 
@@
   geom_point() +
-  labs(x = "Year", y = "Average Percentage of Population", title = "Average Childless Percentage of Population over Time by Marital Status") +
+  geom_smooth(method = "lm", se = FALSE, color = "blue") +
+  labs(x = "Year", y = "Average Percentage per Year") +
   theme_minimal() +
@@
 
+## Between Socio-economic Class, Educational Attainment and Childlessness
+
+While all classes show an increase of childlessness in @fig-nochild-class-edu, the data suggests the most significant rise occurred among the lower class, with their percentage of childless individuals nearly doubling from approximately 17.16% in 1972 to around 34.46% in 2022, marking a substantial increase of approximately 101.04% over the specified period.
+The working class also experienced a significant increase, from around 21.83% in 1972 to approximately 31.44% in 2022, indicating a rise of roughly 43.98%.
+Additionally, the middle class saw a slight increase, from approximately 28.24% in 1972 to about 30.09% in 2022, representing a rise of approximately 6.55%.
+Similarly, the upper class saw an increase, with their percentage of childless individuals climbing from approximately 20.24% in 1972 to about 27.87% in 2022, indicating an increase of roughly 37.71%.
+
+The data suggests that individuals in the lower and working classes exhibit higher percentages of childlessness compared to those in the middle and upper classes.
+This pattern might be attributed to various factors, such as financial constraints, job insecurity, or limited access to childcare resources experienced by individuals in lower socioeconomic classes.
+
+Similarly, data reveals a distinct pattern across educational groups.
+Individuals with a high school diploma consistently exhibit the highest percentage of childlessness throughout the entire period, ranging from around 28.15% in 1972 to 34.56% in 2022.
+Conversely, the percentage of childless individuals with a college and higher degree has decreased from 32.85% in 1972 to 29.84% in 2022.
+The trend for individuals with a with less than a high school education increased from 16.97% in 1972 to 22.90% in 2022.
+
+Thus, the data indicates a nuanced relationship.
+While individuals with higher educational degrees (college+) display a slightly higher percentage of childlessness compared to those with lower degrees, the difference is not as substantial as observed with socioeconomic class.
+This suggests that educational attainment might play a complex role in childbearing decisions, potentially influenced by factors like career aspirations, financial considerations, and personal preferences.
+
 ```{r}
+#| include: true
+#| label: fig-avg-two
+#| fig-cap: Average Childless Percentage per Year in the US (1972-2022)
+#| fig-subcap: [Childless percentage by class, Childless percentage by degree]
+#| layout-ncol: 2
+#| warning: false
+#| message: false
 #| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-avg-class
-#| fig-cap: "Average Percentage Childless over Time by Class"
+#| fig-pos: "t"
 
@@
   geom_point() +
-  labs(x = "Year", y = "Average Percentage of Population", title = "Average Percentage Childless over Time by Class") +
+  geom_smooth(method = "lm", se = FALSE, color = "blue") +
+  labs(x = "Year", y = "Average Percentage per Year") +
   theme_minimal() +
   theme(text = element_text(size = 9))
-
-```
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-avg-gender
-#| fig-cap: "Average Percentage Childless over Time by Gender"
-
-# Group the data by Year and calculate the average of Percentage for each year
-avg_data_sex <- data_long_sex |>
-  group_by(Year) |>
-  summarise(Sum_Percentage = sum(Percentage, na.rm = TRUE)) |>
-  mutate(avg_Percentage = Sum_Percentage / 2)
-
-# Print the average data
-# print(avg_data_sex)
-
-# Plot the data
-ggplot(avg_data_sex, aes(x = Year, y = avg_Percentage)) +
-  geom_line() +
-  geom_point() +
-  labs(x = "Year", y = "Average Percentage of Population", title = "Average Percentage Childless over Time by Gender") +
-  theme_minimal() +
-  theme(text = element_text(size = 9))
-
-```
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-avg-degree
-#| fig-cap: "Average Percentage Childless over Time by Degree"
 
@@
   geom_point() +
-  labs(x = "Year", y = "Average Percentage of Population", title = "Average Percentage Childless over Time by Degree") +
+  geom_smooth(method = "lm", se = FALSE, color = "blue") +
+  labs(x = "Year", y = "Average Percentage per Year") +
   theme_minimal() +
@@
 
+## Overall Rate of Childlessness Over Time
+
 ```{r}
+#| include: true
+#| label: fig-overall
+#| fig-cap: Overall Childless Percentage per Year in the US (1972-2022)
+#| warning: false
+#| message: false
 #| echo: false
-#| include: true
-#| warning: false
-#| label: fig-nochild-avg-degree-class
-#| fig-cap: "Average Percentage Childless over Time by Degree and Class"
+#| fig-pos: "t"
 
-# Merge the two datasets based on the Year column
-merged_data <- merge(avg_data_degree, avg_data_class, by = "Year", all = TRUE, suffixes = c("_degree", "_class"))
+data_long_overall <- pivot_longer(table_nochild_overall, -`Year`, names_to = "year", values_to = "Percentage")
 
-# Calculate the sum of avg_Percentage for each year
-merged_data <- merged_data |>
-  mutate(sum_Percentage = (coalesce(avg_Percentage_degree, 0) + coalesce(avg_Percentage_class, 0))/2)
+data_long_overall <- data_long_overall |>
+select(-`Year`) |>
+rename(`Year` = `year`) |>
+arrange(`Year`) |>
+mutate_all(as.numeric)
 
-# Print the merged and summed data
-# print(merged_data)
+# Print the average data
+# print(data_long_overall)
 
-# Plot the merged and summed data
-ggplot(merged_data, aes(x = Year, y = sum_Percentage)) +
-  geom_line() +
+# Plot the data
+data_long_overall$Year <- as.numeric(data_long_overall$Year)
+data_long_overall$Percentage <- as.numeric(data_long_overall$Percentage)
+
+# Plot the data with points, line, and trend line
+ggplot(data_long_overall, aes(x = Year, y = Percentage)) +
   geom_point() +
-  labs(x = "Year", y = "Sum of Average Percentage of Population", title = "Sum of Average Percentage Childless over Time by Degree and Class") +
+  geom_line() +  # Add a line connecting the points
+  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add a linear trend line
+  labs(x = "Year", y = "Overall Percentage per Year") +
   theme_minimal() +
-  theme(text = element_text(size = 9))
+  theme(text = element_text(size = 9)) +
+  scale_x_continuous(labels = function(x) as.character(x))  # Convert numeric labels to character if needed
+
+```
+
+```{r}
+#| include: false
+# Calculate the correlation coefficient
+correlation <- cor(data_long_overall$Year, data_long_overall$Percentage)
+
+# Print the correlation coefficient
+print(paste("Correlation coefficient between Year and Percentage:", correlation))
 ```
@@
 \newpage
+@fig-overall graphs a trend-line of overall childlessness rate throughout the years.
+With a correlation coefficient between "Year" and "Percentage" of 0.73, it's highly likely that as time progresses, childlessness is on the rise.
+This is evident in the 6.99% increase in childlessness from 1972 to 2022.
+
+To account for the general increase of childlessness from 1972 to 2022, there are several crucial events.
+First, in @fig-overall, the first surge of childlessness occurs from 1986 onwards till 1900 (27.71% to 30.26%, a 2.55% increase).
+This was when the AIDs pandemic occurred, where there was much fear and health concerns of HIV transmission in having sexual intercourse.
+The second surge of childlessness is observable from 2000 onwards till 2006 (27.76% to 28.94%, a 1.18% increase).
+This was the time period marked by fear of terrorism when 9/11 happened, where heightened feelings of vulnerability may have influenced people's decisions of expanding their family.
+The third surge of childlessness from 2008 till 2014 is drastic due to the occurrence of the Great Recession (27.25% to 30.75%, a 3% increase).
+Financial instability then is a contributing factor in determining childlessness.
+The last surge occured from 2018 to 2021 when the COVID-19 was in place (29.69% to 31.67%, a 1.98% increase).
 
@@
 The findings presented in this paper reveal a consistent and concerning trend of rising childlessness across various demographics in the United States.
-As @fig-nochild-age illustrates, the percentage of individuals without children has steadily increased across all age groups over the past 50 years, with the most significant rise observed among younger individuals (18-34 and 35-49).
+As @fig-overall illustrates, the percentage of individuals without children has steadily increased across all age groups over the past 50 years, with the most significant rise observed among younger individuals (18-34 and 35-49).
 
@@
 This paper cannot definitively establish causal relationships, but the data suggests several potential factors contributing to rising childlessness, all supported by existing research.
-These include economic concerns like rising childcare costs, student loan debt, and economic uncertainty.
-Additionally, individuals with higher educational attainment, particularly women, might delay or forgo childbearing to pursue career opportunities.
+These include economic concerns like two global pandemics, rising childcare costs, student loan debt, and economic uncertainty [@voluntaryChild].
+Additionally, individuals with higher educational attainment, particularly women, might delay or forgo childbearing to pursue career opportunities [@effectiveVol].
 Furthermore, shifting social norms related to childbearing, family structures, and career aspirations, especially for women, may be contributing factors.
@@
 A shrinking pool of younger workers could pose challenges to economic growth and strain social security systems.
-Additionally, the rise of childless families could have implications for social support networks and intergenerational relationships.
-While childlessness can be a personal choice for some individuals and may even lead to a fulfilling life, it could also be associated with feelings of isolation and a lack of social support for others.
+Additionally, the rise of childless families could have implications for social support networks and intergenerational relationships [@voluntaryChild].
+While childlessness can be a personal choice for some individuals and may even lead to a fulfilling life, it could also be associated with feelings of isolation and a lack of social support for others [@effectiveVol].
 These potential consequences highlight the need for further research and discussion on how to navigate the social, economic, and individual complexities associated with rising childlessness.
@@
 
-This paper investigated the rising trend of childlessness in the United States, highlighting its prevalence across various demographics.
-By analyzing data from the GSS, it revealed consistent and concerning increases in the percentage of individuals without children.
-While definitive causal explanations remain elusive due to limitations in the analysis, the discussion explored potential contributing factors, including economic considerations, educational aspirations, shifting social norms, and individual preferences.
-Additionally, the potential consequences of this trend, spanning across social, economic, and individual realms, were examined.
+This paper investigates the rising trend of childlessness in the United States, highlighting its prevalence across various demographics.
+By analyzing data from the GSS, it reveals consistent and concerning increases in the percentage of individuals without children.
+While definitive causal explanations remain elusive due to limitations in the analysis, the discussion explores potential contributing factors, including economic considerations, educational aspirations, shifting social norms, and individual preferences.
 
 Moving forward, further research with longitudinal studies and diverse data sources including qualitative interviews, alongside a mixed methods approach, is crucial to gain a comprehensive understanding of childlessness, its nuances, and its multifaceted implications.
-This knowledge can inform future policies and interventions aimed at fostering a supportive environment for individuals and families as they navigate the complex decisions around childbearing in the 21st century.
+This knowledge can inform future policies aimed at fostering a supportive environment for individuals and families as they navigate the complex decisions around childbearing in the 21st century.
 

### 2024-03-18 scripts/02-data_cleaning.R

@@ -20,7 +20,8 @@ input_files <- c(
   "nochild_sex.csv",
   "nochild_higestdegree.csv",
   "nochild_health.csv",
-  "nochild_marital.csv"
+  "nochild_marital.csv",
+  "nochild_overall.csv"
 )
 
 # Remove "nochild_" from input file names and create output file names

### 2024-03-18 outputs/paper/paper.qmd
Context arragment change

@@
 Traditionally viewed as a societal norm, having children now becomes a choice [@voluntaryChild].
-We examine factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they affect childbearing rates over the past 50 years.
+We examine factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they have affected childbearing rates over the past 50 years.
 
@@
 
-Our estimand is the relationship between age, gender, socio-economic class, educational attainment, health status, marital status and the rate of childlessness in the United States from 1972 to 2022.
-Through our analysis, we found that childless rate is consistently on the rise from 1972 to 2022, with correlation coefficient of 0.73 between childless percentage and time (or a 6.99% increase from 1972 to 2022).
-Factors that directly contribute to this rise are higher educational attainment, higher socio-economic class,lower gender disparity, better health status and dissolution of marriages through divorces, separation and being widowed.
+Our estimand is the relationship between age, gender, socioeconomic class, educational attainment, health status, marital status and the rate of childlessness in the United States from 1972 to 2022.
+Through our analysis, we found that the childless rate is consistently on the rise from 1972 to 2022, with a correlation coefficient of 0.73 between the childless percentage and time (or a 6.99% increase from 1972 to 2022).
+Factors that directly contribute to this rise are higher educational attainment, higher socioeconomic class, lower gender disparity, better health status and dissolution of marriages through divorces, separation and being widowed.
 
@@
 For instance, the General Social Survey (GSS) obtains information on marital status through questions such as, "Are you currently married, widowed, divorced, separated, or have you never been married?", with responses categorized into predefined groups for further analysis [@gssCodebook].
-On ther other hand, instead of relying on income sources, socio-economic class is subjectively assessed based on the respondents' answers to this question, "If you were asked to use one of four names for your social class, which would you say you belong in: the lower class, the working class, the middle class, or the upper class?".
-Some are continuous variable, like age.
+On the other hand, instead of relying on income sources, socioeconomic class is subjectively assessed based on the respondents' answers to this question, "If you were asked to use one of four names for your social class, which would you say you belong in: the lower class, the working class, the middle class, or the upper class?".
+Some are continuous variables, like age.
 With respondents' various age groups documented, we can understand the generational disparities in childbearing patterns.
-Systematic and data-driven, these measurement methods ensures that all data faithfully reflects the American population's childbearing tendencies.
+Systematic and data-driven, these measurement methods ensure that all data faithfully reflects the American population's childbearing tendencies.
 
@@
 However, this multimodal survey approach also presents several challenges in collecting respondents' opinions on topics such as childbirth and childbearing tendencies, given the private nature of these subjects.
-With the easy formatting and non-required nature of these surveys, biases like acquiescence bias (only positive responses) might occur because respondents aren't confronted in a face-to-face settings, where many will just opt a neutral response [@webSurvey].
+With the easy formatting and non-required nature of these surveys, biases like acquiescence bias (only positive responses) might occur because respondents aren't confronted in a face-to-face setting, where many will just opt for a neutral response [@webSurvey].
 Therefore, we carefully select various variables to to ensure the authenticity of the public's background on the topic of childlessness, avoiding having the data being too influenced by the changes in surveying methods.
@@
 Thus, there is a consistent rise in childlessness across all age groups in the United States from 1972 to 2022.
-Notably, the most significant rise is observed among younger adults (18-34 and 35-49).
-It is important to acknowledge, however, that the data series has gaps for several years within this timeframe (1979, 1981, 1992, 1995, 1997, 1999, 2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017, and 2019).
+The most significant rise is observed among younger adults (18-34 and 35-49).
+However, it is important to acknowledge that the data series has gaps for several years within this timeframe (1979, 1981, 1992, 1995, 1997, 1999, 2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017, and 2019).
 These missing data points can potentially limit the generalization of the observed trends and hinder a comprehensive understanding of childlessness over time.
@@
 However, females consistently exhibit a lower percentage of childlessness compared to males throughout the entire period.
-While the gap appears to be slightly narrowing over time, females still have a lower percentage by around 8 percentage points in 2022 (approximately 28.32% for females and 36.09% for males).
+While the gap appears to be slightly narrowing over time, females still have a lower percentage by around eight percentage points in 2022 (approximately 28.32% for females and 36.09% for males).
 
@@
 
+From the GSS, we utilize several key variables to enrich our analysis and explore the potential influence of various factors on childbearing decisions:
+
+* **Socioeconomic Class:** Data on socioeconomic class, often measured by income or occupation, will be utilized to understand the impact of economic factors on childbearing choices.
+* **Educational Attainment:** Data on educational attainment, such as the highest degree obtained, will be incorporated to explore the relationship between educational level and childbearing.
+
+### Overview of Socioeconomic Class's Impact on Childlessness
+
+@tbl-nochild-class reveals a potential link between socioeconomic class and childlessness.
+To visualize the data, @fig-nochild-class-edu depicts the gradual increase in the percentage of childless individuals across four social classes (lower class, working class, middle class, and upper class) in the United States from 1972 to 2022.
+The graph reveals an upward trend in childlessness for all classes over the five decades.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-class
+
+# Selected years to show better on the paper
+random_years <- c("Class Identification", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_class_subset <- table_nochild_class |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_class_subset, 
+      caption = "Childless Percentage of Population by Class Identification and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
+People and households are divided into groups based on their socioeconomic status.
+The lower class experiences economic hardship, while the working class earns limited wages and faces job insecurity.
+The middle class has white-collar or professional occupations and enjoys a comfortable lifestyle.
+The upper class has substantial wealth and enjoys a privileged lifestyle with exclusive amenities [@pewresearch2022].
+
+\newpage
+### Overview of Educational Attainment's Impact on Childlessness
+
+Further, @tbl-nochild-degree explores the association between educational attainment and childlessness.
+@fig-nochild-class-edu illustrates this data, presenting a comparative analysis of childlessness across individuals with different levels of educational attainment (less than high school, high school and college degree or higher) in the United States from 1972 to 2022.
+The graph utilizes line trends to depict the percentage of childless individuals within each degree category over the five decades.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-degree
+
+# Selected years to show better on the paper
+random_years <- c("Degree", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_degree_subset <- table_nochild_degree |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_degree_subset, 
+      caption = "Childless Percentage of Population by Degree and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
 ```{r}
@@
 
-From the GSS, we utilize several key variables to enrich our analysis and explore the potential influence of various factors on childbearing decisions:
+## Exploring Health Conditions and Marital Status
 
-* **Socioeconomic Class:** Data on socioeconomic class, often measured by income or occupation, will be utilized to understand the impact of economic factors on childbearing choices.
-* **Educational Attainment:** Data on educational attainment, such as the highest degree obtained, will be incorporated to explore the relationship between educational level and childbearing.
+In addition, additional data sets from NORC [@gss] were also leveraged to understand the context of childlessness in America from 1972-2022.
+@tbl-nochild-health and @tbl-nochild-marital present an interesting connection between childlessness, health conditions and marital status.
 
-### Overview of Socioeconomic Class's Impact on Childlessness
+From the GSS, we utilize several important variables relevant to our analysis:
+
+* **Health Conditions:** This variable, measured by medical check-ups of individuals who have been surveyed, will be used to clarify the context behind the childlessness of those with varying health statuses.
+* **Marital Status:** This variable enables us to see if marriage is a determining factor in childbearing decisions.
+
+### Overview of Health Status's Impact on Childlessness
+
+@tbl-nochild-health uncovers the relationship between health conditions and childlessness.
+Accordingly, the first graph of @fig-nochild-health-marital explores the health differences in people who are childless within the United States over a 50-year period (1972-2022).
+The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (excellent, good, fair, and poor).
 
@@
 #| warning: false
-#| label: tbl-nochild-class
+#| label: tbl-nochild-health
 
 # Selected years to show better on the paper
-random_years <- c("Class Identification", "1972", "1980", "1990", "2000", "2010", "2022")
+random_years <- c("Condition of health", "1972", "1980", "1990", "2000", "2010", "2022")
 
 # Subset the data with only the selected columns
-table_nochild_class_subset <- table_nochild_class |>
+table_nochild_health_subset <- table_nochild_health |>
   select(all_of(random_years))
@@
 # Print the table with formatting
-kable(table_nochild_class_subset, 
-      caption = "Childless Percentage of Population by Class Identification and Selected Years", 
+kable(table_nochild_health_subset, 
+      caption = "Childless Percentage of Population by Health Conditions and Selected Years", 
       align = "c",
@@
 
-@tbl-nochild-class reveals a potential link between socioeconomic class and childlessness.
-To visualize the data, @fig-nochild-class-edu depicts the gradual increase in the percentage of childless individuals across four social classes (lower class, working class, middle class, and upper class) in the United States from 1972 to 2022.
-The graph reveals an upward trend in childlessness for all classes over the five decades.
+### Overview of Marital Status's Impact on Childlessness
 
-People and households are divided into groups based on their socioeconomic status.
-The lower class experiences economic hardship, while the working class earns limited wages and faces job insecurity.
-The middle class has white-collar or professional occupations and enjoys a comfortable lifestyle.
-The upper class has substantial wealth and enjoys a privileged lifestyle with exclusive amenities [@pewresearch2022].
-
-\newpage
-
-### Overview of Educational Attainment's Impact on Childlessness
-
-Further, @tbl-nochild-degree explores the association between educational attainment and childlessness.
-@fig-nochild-class-edu illustrates this data, presenting a comparative analysis of childlessness across individuals with different levels of educational attainment (less than high school, high school and college degree or higher) in the United States from 1972 to 2022.
-The graph utilizes line trends to depict the percentage of childless individuals within each degree category over the five decades.
+@tbl-nochild-marital explores people's marital status and childlessness.
+@fig-nochild-health-marital then illustrates this data, exploring the marital status of people who are childless over a span of 50 years in the United States (1972-2022).
+The graph contains trend lines of varying marital statuses-divorced, married, never married, separated and widowed-and how each has changed over the years in terms of childbearing status.
 
@@
 #| warning: false
-#| label: tbl-nochild-degree
+#| label: tbl-nochild-marital
 
 # Selected years to show better on the paper
-random_years <- c("Degree", "1972", "1980", "1990", "2000", "2010", "2022")
+random_years <- c("Marital status", "1972", "1980", "1990", "2000", "2010", "2022")
 
 # Subset the data with only the selected columns
-table_nochild_degree_subset <- table_nochild_degree |>
+table_nochild_marital_subset <- table_nochild_marital |>
   select(all_of(random_years))
@@
 # Print the table with formatting
-kable(table_nochild_degree_subset, 
-      caption = "Childless Percentage of Population by Degree and Selected Years", 
+kable(table_nochild_marital_subset, 
+      caption = "Childless Percentage of Population by Marital Status and Selected Years", 
       align = "c",
@@
 ```
-
-## Exploring Health Conditions and Marital Status
-
-In addition, additional data sets from NORC [@gss] were also leveraged to understand the context of childlessness in America from 1972-2022.
-@tbl-nochild-health and @tbl-nochild-marital present an interesting connection between childlessness and health conditions and marital status.
-
-From the GSS, we utilize several important variables relevant to our analysis:
-
-* Health Conditions: This variable, measured by medical check-ups of individuals who have been surveyed, will be used to clarify the context behind the childlessness of those with varying health statuses.
-* Marital Status: This variable enables us to see if marriage is a determining factor in childbearing decisions.
 
@@
 
-\newpage
-### Overview of Health Status's Impact on Childlessness
-
-@tbl-nochild-health uncovers the relationship between health conditions and childlessness.
-Accordingly, the first graph of @fig-nochild-health-marital explores the health differences in people who are childless within the United States over a 50-year period (1972-2022).
-The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (excellent, good, fair, and poor).
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: tbl-nochild-health
-
-# Selected years to show better on the paper
-random_years <- c("Condition of health", "1972", "1980", "1990", "2000", "2010", "2022")
-
-# Subset the data with only the selected columns
-table_nochild_health_subset <- table_nochild_health |>
-  select(all_of(random_years))
-
-# Print the table with formatting
-kable(table_nochild_health_subset, 
-      caption = "Childless Percentage of Population by Health Conditions and Selected Years", 
-      align = "c",
-      col.names = random_years,
-      row.names = FALSE) |>
-  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
-
-```
-
-### Overview of Marital Status's Impact on Childlessness
-
-@tbl-nochild-marital explores people's marital status and childlessness.
-@fig-nochild-health-marital then illustrates this data, exploring the marital status of people who are childless over a span of 50 years in the United States (1972-2022).
-The graph contains trend lines of varying marital statuses-Divorced, Married, Never married, Separated and Widowed-and how each has changed over the years in terms of childbearing status.
-
-\newpage
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: tbl-nochild-marital
-
-# Selected years to show better on the paper
-random_years <- c("Marital status", "1972", "1980", "1990", "2000", "2010", "2022")
-
-# Subset the data with only the selected columns
-table_nochild_marital_subset <- table_nochild_marital |>
-  select(all_of(random_years))
-
-# Print the table with formatting
-kable(table_nochild_marital_subset, 
-      caption = "Childless Percentage of Population by Marital Status and Selected Years", 
-      align = "c",
-      col.names = random_years,
-      row.names = FALSE) |>
-  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
-
-```
-
 ## Data Limitations
@@
 
-Survey data may be susceptible to measurement errors due to various factors like social desirability bias, memory lapses, or misinterpretations of questions.
+Survey data may be susceptible to measurement errors due to various factors, such as social desirability bias, memory lapses, or misinterpretations of questions.
 While the GSS implements measures to minimize these errors, they cannot be entirely eliminated.
@@
 However, people with poor health, once having the lowest rate of childlessness (at 12.27% in 1972), now increases to 20.40% in 2022, which translates to a substantial 8.13% increase.
-Overall, on average, across all health levels from 1972-2022, @fig-avg-one shows that percentage of childlessness increases at about 6.836%.
+Overall, on average, across all health levels from 1972-2022, @fig-avg-one shows that the percentage of childlessness increases at about 6.836%.
 
@@
 For married people though, this childlessness percentage also decreases, from 10.51% in 1972 to 13.97% in 2022.
-Overall, on average, across all marital status from 1972-2022, @fig-avg-one shows that percentage of childlessness decreases at about 7.39%.
+Overall, on average, across all marital statuses from 1972-2022, @fig-avg-one shows that the percentage of childlessness decreases to about 7.39%.
 
@@
 
-## Between Socio-economic Class, Educational Attainment and Childlessness
+## Between Socioeconomic Class, Educational Attainment and Childlessness
 
-While all classes show an increase of childlessness in @fig-nochild-class-edu, the data suggests the most significant rise occurred among the lower class, with their percentage of childless individuals nearly doubling from approximately 17.16% in 1972 to around 34.46% in 2022, marking a substantial increase of approximately 101.04% over the specified period.
+While all classes show an increase in childlessness in @fig-nochild-class-edu, the data suggests the most significant rise occurred among the lower class, with their percentage of childless individuals nearly doubling from approximately 17.16% in 1972 to around 34.46% in 2022, marking a substantial increase of approximately 101.04% over the specified period.
 The working class also experienced a significant increase, from around 21.83% in 1972 to approximately 31.44% in 2022, indicating a rise of roughly 43.98%.
@@
 Conversely, the percentage of childless individuals with a college and higher degree has decreased from 32.85% in 1972 to 29.84% in 2022.
-The trend for individuals with a with less than a high school education increased from 16.97% in 1972 to 22.90% in 2022.
+The trend for individuals with less than a high school education increased from 16.97% in 1972 to 22.90% in 2022.
 
@@
 
+@fig-overall graphs a trend line of the overall childlessness rate throughout the years.
+With a correlation coefficient between "Year" and "Percentage" of 0.73, it's highly likely that as time progresses, childlessness is on the rise.
+This is evident in the 6.99% increase in childlessness from 1972 to 2022.
+
+To account for the general increase in childlessness from 1972 to 2022, there are several crucial events.
+First, in @fig-overall, the first surge of childlessness occurs from 1986 onwards till 1900 (27.71% to 30.26%, a 2.55% increase).
+This was when the AIDs pandemic occurred, when there was much fear and health concerns about HIV transmission in having sexual intercourse.
+The second surge of childlessness is observable from 2000 onwards till 2006 (27.76% to 28.94%, a 1.18% increase).
+This was the time period marked by fear of terrorism when 9/11 happened, where heightened feelings of vulnerability may have influenced people's decisions of expanding their family.
+The third surge of childlessness from 2008 till 2014 is drastic due to the occurrence of the Great Recession (27.25% to 30.75%, a 3% increase).
+Financial instability then is a contributing factor in determining childlessness.
+The last surge occurred from 2018 to 2021 when COVID-19 was in place (29.69% to 31.67%, a 1.98% increase).
+
 ```{r}
@@
 ```
-
-\newpage
-@fig-overall graphs a trend-line of overall childlessness rate throughout the years.
-With a correlation coefficient between "Year" and "Percentage" of 0.73, it's highly likely that as time progresses, childlessness is on the rise.
-This is evident in the 6.99% increase in childlessness from 1972 to 2022.
-
-To account for the general increase of childlessness from 1972 to 2022, there are several crucial events.
-First, in @fig-overall, the first surge of childlessness occurs from 1986 onwards till 1900 (27.71% to 30.26%, a 2.55% increase).
-This was when the AIDs pandemic occurred, where there was much fear and health concerns of HIV transmission in having sexual intercourse.
-The second surge of childlessness is observable from 2000 onwards till 2006 (27.76% to 28.94%, a 1.18% increase).
-This was the time period marked by fear of terrorism when 9/11 happened, where heightened feelings of vulnerability may have influenced people's decisions of expanding their family.
-The third surge of childlessness from 2008 till 2014 is drastic due to the occurrence of the Great Recession (27.25% to 30.75%, a 3% increase).
-Financial instability then is a contributing factor in determining childlessness.
-The last surge occured from 2018 to 2021 when the COVID-19 was in place (29.69% to 31.67%, a 1.98% increase).
 

### 2024-03-18 outputs/paper/paper.qmd
update new changes to introduction

--- 2024-03-18 outputs/paper/paper.qmd
@@
 The American family landscape is undergoing a significant transformation, marked by a growing trend of childlessness.
-Traditionally viewed as a societal norm, having children now becomes a choice [@voluntaryChild].
-We examine factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they affect childbearing rates over the past 50 years.
+Traditionally viewed as a social norm, having children now becomes a choice [@voluntaryChild].
+We examine factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they affect childbearing rates from 1972 to 2022.
+Existing research suggests a complex interplay of factors that contribute to childlessness, including economic considerations, rising educational attainment, shifting social norms related to career and family life, and individual preferences [@voluntaryChild].
 
-These helps us understand not only the prevalence of childlessness within different segments of the population but also the potential drivers behind this trend.
-Existing research suggests a complex interplay of factors, including economic considerations, rising educational attainment, changing social norms related to career and family life, and individual preferences [@voluntaryChild].
-
-We explore the impact of these factors on the rise of childlessness in America.
 By analyzing the data, we seek to answer crucial questions: What are some important factors contributing to this trend?
@@
 Our estimand is the relationship between age, gender, socio-economic class, educational attainment, health status, marital status and the rate of childlessness in the United States from 1972 to 2022.
-Through our analysis, we found that childless rate is consistently on the rise from 1972 to 2022, with correlation coefficient of 0.73 between childless percentage and time (or a 6.99% increase from 1972 to 2022).
-Factors that directly contribute to this rise are higher educational attainment, higher socio-economic class,lower gender disparity, better health status and dissolution of marriages through divorces, separation and being widowed.
+Through our analysis, we found that childless rate is consistently on the rise from 1972 to 2022, with correlation coefficient of 0.73 between childless percentage and time (or a 6.99% increase over the last 50 years).
+Factors that directly contribute to this rise are higher educational attainment, higher socio-economic class, lower gender disparity, better health status and dissolution of marriages through divorces, separation and being widowed.
+In addition, global events like pandemics and recession also heighten the rate of childlessness from 1972-2022, including the AIDs pandemic, 9/11, 2008 Recession and the COVID-19 pandemic.
+During the 2008 Recession, from 2008 to 2014, the rate of childlessness increases the most, at 3%.
 

### 2024-03-18 outputs/paper/paper.qmd
GSS questions added

--- 2024-03-18 outputs/paper/paper.qmd
@@
 # Appendix {#sec-appendix}
-The supplementary survey is accessible through this link: https://forms.gle/DsFgJCKPeCtdqZKL6
 
-## Survey Preamble
+## Our Survey for Childlessness Study
+
+The supplementary survey is accessible through this [Google Form Link](https://forms.gle/DsFgJCKPeCtdqZKL6)
+
+### Survey Preamble
 The General Social Survey (GSS) collects data on social trends to help track changes in Canadians' living conditions and well-being over time.
@@
 
-## Survey Questions for Childlessness
+### Survey Questions
 
@@
 
+### Survey Completion Message
+
 **Thank you for your participation**
@@
 
+## Survey Questions from GSS
+
+1. RESPONDENT'S AGE: (short answer text)
+
+2. RESPONDENT'S SEX
+
+- Female
+- Male
+
+3. Are you currently--married, widowed, divorced, separated, or have you never been married? (short answer text)
+
+4. RESPONDENT'S DEGREE
+
+- Graduate
+- Bachelor's
+- Associate/junior college
+- High school
+- Less than high school
+
+5. How many children have you ever had?  Please count all that were born alive at any time (including any you had from a previous marriage).
+
+- 0
+- 1
+- 2
+- 3
+- 4
+- 5
+- 6
+- 7
+- 8 or more
+
+6. If you were asked to use one of four names for your social class, which would you say you belong in: the lower class, the working class, the middle class, or the upper class? (short answer text)
+
+7. Would you say your own health, in general, is excellent, good, fair, or  poor? (short answer text)
+
 \newpage

### 2024-03-18 outputs/paper/paper.qmd
re-format paper

--- 2024-03-18 outputs/paper/paper.qmd
@@
 # Introduction {#sec-intro}
+
 The American family landscape is undergoing a significant transformation, marked by a growing trend of childlessness.
-Traditionally viewed as a societal norm, having children now becomes a choice [@voluntaryChild].
-We examine factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they have affected childbearing rates over the past 50 years.
+Traditionally viewed as a social norm, having children now becomes a choice [@voluntaryChild].
+We examine factors such as age, gender, education, socioeconomic class, health conditions and marital status, and how they affect childbearing rates from 1972 to 2022.
+Existing research suggests a complex interplay of factors that contribute to childlessness, including economic considerations, rising educational attainment, shifting social norms related to career and family life, and individual preferences [@voluntaryChild].
 
-These helps us understand not only the prevalence of childlessness within different segments of the population but also the potential drivers behind this trend.
-Existing research suggests a complex interplay of factors, including economic considerations, rising educational attainment, changing social norms related to career and family life, and individual preferences [@voluntaryChild].
-
-We explore the impact of these factors on the rise of childlessness in America.
 By analyzing the data, we seek to answer crucial questions: What are some important factors contributing to this trend?
@@
 
-Our estimand is the relationship between age, gender, socioeconomic class, educational attainment, health status, marital status and the rate of childlessness in the United States from 1972 to 2022.
-Through our analysis, we found that the childless rate is consistently on the rise from 1972 to 2022, with a correlation coefficient of 0.73 between the childless percentage and time (or a 6.99% increase from 1972 to 2022).
-Factors that directly contribute to this rise are higher educational attainment, higher socioeconomic class, lower gender disparity, better health status and dissolution of marriages through divorces, separation and being widowed.
+Thus, our estimand is the relationship between age, gender, socio-economic class, educational attainment, health status, marital status and the rate of childlessness in the United States from 1972 to 2022.
+Through our analysis, we found that childless rate is consistently on the rise from 1972 to 2022, with correlation coefficient of 0.73 between childless percentage and time (or a 6.99% increase over the last 50 years).
+Factors that directly contribute to this rise are higher educational attainment, higher socio-economic class, lower gender disparity, better health status and dissolution of marriages through divorces, separation and being widowed.
+In addition, global events like pandemics and recession also heighten the rate of childlessness from 1972-2022, including the AIDs pandemic, 9/11, 2008 Recession and the COVID-19 pandemic.
+During the 2008 Recession, from 2008 to 2014, the rate of childlessness increases the most, at 3%.
 
@@
 ## Exploring Gender Differences
-It is crucial to acknowledge potential gender disparities in childbearing experiences.
 A separate analysis examining childlessness by gender (presented in @tbl-nochild-sex) aims to provide a comprehensive understanding by considering the unique experiences of men and women regarding childlessness.
@@
 ## Exploring Socioeconomic Class and Educational Attainment
-To gain a deeper understanding of the factors influencing these trends, we look into additional data sets from NORC [@gss] examining the relationship between childlessness and socioeconomic class (@tbl-nochild-class) and educational attainment (@tbl-nochild-degree).
+We also look into additional data sets from NORC [@gss] examining the relationship between childlessness and socioeconomic class (@tbl-nochild-class) and educational attainment (@tbl-nochild-degree).
 
@@
 * **Educational Attainment:** Data on educational attainment, such as the highest degree obtained, will be incorporated to explore the relationship between educational level and childbearing.
+
+```{r}
+#| include: true
+#| label: fig-nochild-class-edu
+#| fig-cap: Childless Percentage in the US by Class and Educational Attainment (1972-2020)
+#| fig-subcap: [Childless percentage by class, Childless percentage by education]
+#| layout-ncol: 2
+#| warning: false
+#| message: false
+#| echo: false
+#| fig-pos: "t"
+
+# Pivot the data from wide to long format
+data_long_class <- pivot_longer(table_nochild_class, -`Class Identification`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_class$Year <- as.integer(data_long_class$Year)
+
+# Plot the data
+ggplot(data_long_class, aes(x = Year, y = Percentage, color = `Class Identification`)) +
+   geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Class", color = "Breakdown Category") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+# Pivot the data from wide to long format
+data_long_degree <- pivot_longer(table_nochild_degree, -`Degree`, names_to = "Year", values_to = "Percentage")
+
+# Convert Year column to integer
+data_long_degree$Year <- as.integer(data_long_degree$Year)
+
+# Plot the data
+ggplot(data_long_degree, aes(x = Year, y = Percentage, color = `Degree`)) +
+  geom_line() +
+  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Degree", color = "Degree") +
+  theme_minimal() +
+  theme(text = element_text(size = 10))
+
+```
 
@@
 
-```{r}
-#| include: true
-#| label: fig-nochild-class-edu
-#| fig-cap: Childless Percentage in the US by Class and Educational Attainment (1972-2020)
-#| fig-subcap: [Childless percentage by class, Childless percentage by education]
-#| layout-ncol: 2
-#| warning: false
-#| message: false
-#| echo: false
-#| fig-pos: "t"
-
-# Pivot the data from wide to long format
-data_long_class <- pivot_longer(table_nochild_class, -`Class Identification`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_class$Year <- as.integer(data_long_class$Year)
-
-# Plot the data
-ggplot(data_long_class, aes(x = Year, y = Percentage, color = `Class Identification`)) +
-   geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Class", color = "Breakdown Category") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-# Pivot the data from wide to long format
-data_long_degree <- pivot_longer(table_nochild_degree, -`Degree`, names_to = "Year", values_to = "Percentage")
-
-# Convert Year column to integer
-data_long_degree$Year <- as.integer(data_long_degree$Year)
-
-# Plot the data
-ggplot(data_long_degree, aes(x = Year, y = Percentage, color = `Degree`)) +
-  geom_line() +
-  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Degree", color = "Degree") +
-  theme_minimal() +
-  theme(text = element_text(size = 10))
-
-```
-
 ## Exploring Health Conditions and Marital Status
 
-In addition, additional data sets from NORC [@gss] were also leveraged to understand the context of childlessness in America from 1972-2022.
+We also use additional data sets from NORC [@gss] to understand the context of childlessness in America from 1972-2022.
 @tbl-nochild-health and @tbl-nochild-marital present an interesting connection between childlessness, health conditions and marital status.
@@
 * **Marital Status:** This variable enables us to see if marriage is a determining factor in childbearing decisions.
-
-### Overview of Health Status's Impact on Childlessness
-
-@tbl-nochild-health uncovers the relationship between health conditions and childlessness.
-Accordingly, the first graph of @fig-nochild-health-marital explores the health differences in people who are childless within the United States over a 50-year period (1972-2022).
-The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (excellent, good, fair, and poor).
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: tbl-nochild-health
-
-# Selected years to show better on the paper
-random_years <- c("Condition of health", "1972", "1980", "1990", "2000", "2010", "2022")
-
-# Subset the data with only the selected columns
-table_nochild_health_subset <- table_nochild_health |>
-  select(all_of(random_years))
-
-# Print the table with formatting
-kable(table_nochild_health_subset, 
-      caption = "Childless Percentage of Population by Health Conditions and Selected Years", 
-      align = "c",
-      col.names = random_years,
-      row.names = FALSE) |>
-  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
-
-```
-
-### Overview of Marital Status's Impact on Childlessness
-
-@tbl-nochild-marital explores people's marital status and childlessness.
-@fig-nochild-health-marital then illustrates this data, exploring the marital status of people who are childless over a span of 50 years in the United States (1972-2022).
-The graph contains trend lines of varying marital statuses-divorced, married, never married, separated and widowed-and how each has changed over the years in terms of childbearing status.
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: tbl-nochild-marital
-
-# Selected years to show better on the paper
-random_years <- c("Marital status", "1972", "1980", "1990", "2000", "2010", "2022")
-
-# Subset the data with only the selected columns
-table_nochild_marital_subset <- table_nochild_marital |>
-  select(all_of(random_years))
-
-# Print the table with formatting
-kable(table_nochild_marital_subset, 
-      caption = "Childless Percentage of Population by Marital Status and Selected Years", 
-      align = "c",
-      col.names = random_years,
-      row.names = FALSE) |>
-  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
-
-```
 
@@
   theme(text = element_text(size = 10))
+
+```
+
+\newpage
+### Overview of Health Status's Impact on Childlessness
+
+@tbl-nochild-health uncovers the relationship between health conditions and childlessness.
+Accordingly, the first graph of @fig-nochild-health-marital explores the health differences in people who are childless within the United States over a 50-year period (1972-2022).
+The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (excellent, good, fair, and poor).
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-health
+
+# Selected years to show better on the paper
+random_years <- c("Condition of health", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_health_subset <- table_nochild_health |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_health_subset, 
+      caption = "Childless Percentage of Population by Health Conditions and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
+### Overview of Marital Status's Impact on Childlessness
+
+@tbl-nochild-marital explores people's marital status and childlessness.
+@fig-nochild-health-marital then illustrates this data, exploring the marital status of people who are childless over a span of 50 years in the United States (1972-2022).
+The graph contains trend lines of varying marital statuses-divorced, married, never married, separated and widowed-and how each has changed over the years in terms of childbearing status.
+
+\newpage
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-marital
+
+# Selected years to show better on the paper
+random_years <- c("Marital status", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_marital_subset <- table_nochild_marital |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_marital_subset, 
+      caption = "Childless Percentage of Population by Marital Status and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
 
@@
 
-## Societal Implications of Rising Childlessness
+## Social and Ethical Implications of Rising Childlessness
 
@@
 We are conducting a study to investigate the reasons behind childlessness in different groups of people.
-To do so, we will analyze various data sets, including information on age, gender, education, socioeconomic status, health conditions, and marital status.
-By examining these factors, we hope to gain insight into how they impact rates of childbearing.
+To do so, we will expand upon the GSS's existing questions and various data sets, including information on age, gender, education, socioeconomic status, health conditions, and marital status.
+By adding additional questions on these factors, we hope to gain more insight into how they impact childbearing rates.
 

### 2024-03-18 outputs/paper/paper.qmd
finalize paper

@@
   geom_smooth(method = "loess", se = FALSE, linetype = "dashed", size = 1.2) +  # Add running average trend line
-  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Class", color = "Breakdown Category") +
+  labs(x = "Year", y = "Percentage of Population", title = "Childless Percentage of Population over Time by Class", color = "Class") +
   theme_minimal() +
@@
 @fig-nochild-class-edu illustrates this data, presenting a comparative analysis of childlessness across individuals with different levels of educational attainment (less than high school, high school and college degree or higher) in the United States from 1972 to 2022.
-The graph utilizes line trends to depict the percentage of childless individuals within each degree category over the five decades.
+The graph utilizes trend lines to depict the percentage of childless individuals within each degree category over the five decades.
 
@@
 # Result {#sec-result}
+
+Upon assessing the available data, we further extrapolate on them here by calculating the average rate of childlessness in each year for each of these four critical factors: health conditions, marital status, educational attainment and socio-economic class.
+At the end, @fig-overall compounds on these factors as a graph that shows the overall rate of childlessness in each year from 1972 to 2022.
 

### 2024-03-18 outputs/paper/paper.qmd
Last edit

@@
 
-Thus, our estimand is the relationship between age, gender, socio-economic class, educational attainment, health status, marital status and the rate of childlessness in the United States from 1972 to 2022.
-Through our analysis, we found that childless rate is consistently on the rise from 1972 to 2022, with correlation coefficient of 0.73 between childless percentage and time (or a 6.99% increase over the last 50 years).
-Factors that directly contribute to this rise are higher educational attainment, higher socio-economic class, lower gender disparity, better health status and dissolution of marriages through divorces, separation and being widowed.
-In addition, global events like pandemics and recession also heighten the rate of childlessness from 1972-2022, including the AIDs pandemic, 9/11, 2008 Recession and the COVID-19 pandemic.
+Thus, our estimand is the relationship between age, gender, socioeconomic class, educational attainment, health status, marital status, and the rate of childlessness in the United States from 1972 to 2022.
+Through our analysis, we found that the childless rate is consistently on the rise from 1972 to 2022, with a correlation coefficient of 0.73 between the childless percentage and time (or a 6.99% increase over the last 50 years).
+Factors that directly contribute to this rise are higher educational attainment, higher socioeconomic class, lower gender disparity, better health status and dissolution of marriages through divorces, separation and being widowed.
+In addition, global events like pandemics and recessions also heighten the rate of childlessness from 1972-2022, including the AIDs pandemic, 9/11, 2008 Recession and the COVID-19 pandemic.
 During the 2008 Recession, from 2008 to 2014, the rate of childlessness increases the most, at 3%.
@@
 With the easy formatting and non-required nature of these surveys, biases like acquiescence bias (only positive responses) might occur because respondents aren't confronted in a face-to-face setting, where many will just opt for a neutral response [@webSurvey].
-Therefore, we carefully select various variables to to ensure the authenticity of the public's background on the topic of childlessness, avoiding having the data being too influenced by the changes in surveying methods.
+Therefore, we carefully select various variables to ensure the authenticity of the public's background on the topic of childlessness, avoiding having the data being too influenced by the changes in surveying methods.
 
@@
 
+### Overview of Socioeconomic Class's Impact on Childlessness
+
+@tbl-nochild-class reveals a potential link between socioeconomic class and childlessness.
+To visualize the data, @fig-nochild-class-edu depicts the gradual increase in the percentage of childless individuals across four social classes (lower class, working class, middle class, and upper class) in the United States from 1972 to 2022.
+The graph reveals an upward trend in childlessness for all classes over the five decades.
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-class
+
+# Selected years to show better on the paper
+random_years <- c("Class Identification", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_class_subset <- table_nochild_class |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_class_subset, 
+      caption = "Childless Percentage of Population by Class Identification and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
 ```{r}
@@
 #| echo: false
-#| fig-pos: "t"
 
@@
 
-### Overview of Socioeconomic Class's Impact on Childlessness
-
-@tbl-nochild-class reveals a potential link between socioeconomic class and childlessness.
-To visualize the data, @fig-nochild-class-edu depicts the gradual increase in the percentage of childless individuals across four social classes (lower class, working class, middle class, and upper class) in the United States from 1972 to 2022.
-The graph reveals an upward trend in childlessness for all classes over the five decades.
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: tbl-nochild-class
-
-# Selected years to show better on the paper
-random_years <- c("Class Identification", "1972", "1980", "1990", "2000", "2010", "2022")
-
-# Subset the data with only the selected columns
-table_nochild_class_subset <- table_nochild_class |>
-  select(all_of(random_years))
-
-# Print the table with formatting
-kable(table_nochild_class_subset, 
-      caption = "Childless Percentage of Population by Class Identification and Selected Years", 
-      align = "c",
-      col.names = random_years,
-      row.names = FALSE) |>
-  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
-
-```
-
 People and households are divided into groups based on their socioeconomic status.
@@
 
-\newpage
 ### Overview of Educational Attainment's Impact on Childlessness
@@
 
+\newpage
+### Overview of Health Status's Impact on Childlessness
+
+@tbl-nochild-health uncovers the relationship between health conditions and childlessness.
+Accordingly, the first graph of @fig-nochild-health-marital explores the health differences in people who are childless within the United States over a 50-year period (1972-2022).
+The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (excellent, good, fair, and poor).
+
+```{r}
+#| echo: false
+#| include: true
+#| warning: false
+#| label: tbl-nochild-health
+
+# Selected years to show better on the paper
+random_years <- c("Condition of health", "1972", "1980", "1990", "2000", "2010", "2022")
+
+# Subset the data with only the selected columns
+table_nochild_health_subset <- table_nochild_health |>
+  select(all_of(random_years))
+
+# Print the table with formatting
+kable(table_nochild_health_subset, 
+      caption = "Childless Percentage of Population by Health Conditions and Selected Years", 
+      align = "c",
+      col.names = random_years,
+      row.names = FALSE) |>
+  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
+
+```
+
 ```{r}
@@
   theme(text = element_text(size = 10))
-
-```
-
-\newpage
-### Overview of Health Status's Impact on Childlessness
-
-@tbl-nochild-health uncovers the relationship between health conditions and childlessness.
-Accordingly, the first graph of @fig-nochild-health-marital explores the health differences in people who are childless within the United States over a 50-year period (1972-2022).
-The graph utilizes trend lines to depict the percentage of childless individuals categorized by health conditions (excellent, good, fair, and poor).
-
-```{r}
-#| echo: false
-#| include: true
-#| warning: false
-#| label: tbl-nochild-health
-
-# Selected years to show better on the paper
-random_years <- c("Condition of health", "1972", "1980", "1990", "2000", "2010", "2022")
-
-# Subset the data with only the selected columns
-table_nochild_health_subset <- table_nochild_health |>
-  select(all_of(random_years))
-
-# Print the table with formatting
-kable(table_nochild_health_subset, 
-      caption = "Childless Percentage of Population by Health Conditions and Selected Years", 
-      align = "c",
-      col.names = random_years,
-      row.names = FALSE) |>
-  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
 
@@
 
-Upon assessing the available data, we further extrapolate on them here by calculating the average rate of childlessness in each year for each of these four critical factors: health conditions, marital status, educational attainment and socio-economic class.
+Upon assessing the available data, we further extrapolate on them here by calculating the average rate of childlessness in each year for each of these four critical factors: health conditions, marital status, educational attainment and socioeconomic class.
 At the end, @fig-overall compounds on these factors as a graph that shows the overall rate of childlessness in each year from 1972 to 2022.
@@
 
-(@) If yes, How many children have you ever had? Please count all that were born alive at any time (including any you had from a previous marriage).
+(@) If yes, how many children have you ever had? Please count all that were born alive at any time (including any you had from a previous marriage).
 (short answer text)
@@
 
-3. Are you currently--married, widowed, divorced, separated, or have you never been married? (short answer text)
+3. Are you currently--married, widowed, divorced, separated, or have you never been married?
+
+- Married
+- Widowed
+- Divorced
+- Separated
+- Never been married
 
@@
 
-6. If you were asked to use one of four names for your social class, which would you say you belong in: the lower class, the working class, the middle class, or the upper class? (short answer text)
+6. If you were asked to use one of four names for your social class, which would you say you belong in: the lower class, the working class, the middle class, or the upper class?
 
-7. Would you say your own health, in general, is excellent, good, fair, or  poor? (short answer text)
+- Lower class
+- Working class
+- Middle class
+- Upper class
+
+7. Would you say your own health, in general, is excellent, good, fair, or poor?
+
+- Excellent
+- Good
+- Fair
+- Poor
 

### 2024-03-19 outputs/paper/paper.qmd
update paper's intro

--- 2024-03-18 outputs/paper/paper.qmd
@@
 
-To further understand why these causes explain rising childlessness, our paper starts with a brief dissection of the nature of childlessness in America and the contributing factors in [Introduction].
+To further understand why these causes explain rising childlessness, in [Introduction], our paper briefly discusses the nature of childlessness in America and its contributing factors.
 Subsequently, in [Data] and [Result], we talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts.

### 2024-03-20 outputs/paper/paper.qmd
update paper

@@
 Overall, on average, across all health levels from 1972-2022, @fig-avg-one shows that the percentage of childlessness increases at about 6.836%.
+Observable from the graph, health as a factor is positively correlated to childlessness.
 
@@
 Overall, on average, across all marital statuses from 1972-2022, @fig-avg-one shows that the percentage of childlessness decreases to about 7.39%.
+Observable from the graph, marital status as a factor is negatively correlated to childlessness.
 
@@
 Similarly, the upper class saw an increase, with their percentage of childless individuals climbing from approximately 20.24% in 1972 to about 27.87% in 2022, indicating an increase of roughly 37.71%.
+Overall, on average, across all socio-economic clases from 1972-2022, @fig-avg-two shows that the percentage of childlessness increases at about 9.21%.
+Observable from the graph, class as a factor is positively correlated to childlessness.
 
@@
 
-Similarly, data reveals a distinct pattern across educational groups.
+Similarly, @fig-nochild-class-edu reveals a distinct pattern across educational groups.
 Individuals with a high school diploma consistently exhibit the highest percentage of childlessness throughout the entire period, ranging from around 28.15% in 1972 to 34.56% in 2022.
@@
 The trend for individuals with less than a high school education increased from 16.97% in 1972 to 22.90% in 2022.
+Overall, on average, across all educational levels from 1972-2022, @fig-avg-two shows that the percentage of childlessness increases at about 3.11%.
+Observable from the graph, educational attainment as a factor is slightly positively correlated to childlessness.
 
@@
 
-```
-
-```{r}
-#| include: false
-# Calculate the correlation coefficient
-correlation <- cor(data_long_overall$Year, data_long_overall$Percentage)
-
-# Print the correlation coefficient
-print(paste("Correlation coefficient between Year and Percentage:", correlation))
 ```
